---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  fig.retina = 2,
  message   = FALSE,
  warning   = FALSE
)
if (identical(Sys.getenv("CSDG_README_DOCUMENTATION_ONLY"), "true")) {
  knitr::opts_chunk$set(eval = FALSE)
}
```

# mlr3autoiml

`mlr3autoiml` helps researchers decide whether the evidence supports a conclusion that they draw from an explanation
of a machine learning model, such as "the model relies mainly on self-rated health." It implements
**Claim-Scoped Diagnostic Gates (CSDG)** for the **mlr3 ecosystem**. An analysis runs in its usual order:

1. **Fit** learners and a model without predictors on the same cross-validation splits with `mlr3::benchmark()`
   (`csdg_fit()`).
2. **Look** at held-out permutation feature importance (PFI), also grouped and conditional (`csdg_importance()`),
   or at accumulated local effects (`csdg_effect()`).
3. **State a claim** that the result suggests (`claim_relies_mainly()`, `claim_top_k()`, `claim_order()`,
   `claim_direction()`).
4. **Check it**: the package derives the properties that the claim requires from the 12 gates of the article and
   computes the default checks (`csdg_check()`).
5. **Assess it** with a fixed decision rule (`csdg_assess()`) and test it on new data with the same criteria
   (`csdg_confirm()`).

The package writes the claim, its scope, the criteria, and the observations from the computations. Properties that
cannot be computed, such as validity evidence for a construct, are entered with `csdg_judge()`. Favorable results
never offset a contradicted property, and no aggregate score is computed. Claims, scopes, and evidence can also be
recorded by hand (`csdg_claim()`, `csdg_evidence_record()`, `csdg_adjudicate_claim()`).

| Term | Values |
|---|---|
| Scope elements | quantity, model, procedure, data, meaning (model description, population claim, causal claim), use |
| Evidence roles | required property, established counterevidence, unresolved threat, context |
| Status of a required property | supported, contradicted, open |
| Assessment of a claim | not met if a required property is contradicted; otherwise unresolved if one is open; otherwise met |
| Decision | a met claim is retained; a claim that is not met or unresolved is revised or withheld |

## Installation

Install the current version from GitHub:

```{r install, eval = FALSE}
remotes::install_github("coorsaa/mlr3autoiml", build_vignettes = TRUE)
library(mlr3autoiml)
```

Released versions are tagged, for example `remotes::install_github("coorsaa/mlr3autoiml@v0.2.0")`.

## Quick start

A simulated survey of 500 respondents has 12 items and a binary outcome. Two scale scores, `low_mood` and
`fatigue`, predict it strongly; `worry_a` and `worry_b` are two wordings of one worry item; three items have weak
effects and five none.

```{r quick-start-data}
library(mlr3)
library(mlr3learners)
library(mlr3autoiml)
library(data.table)

simulate_survey = function(n, seed) {
  set.seed(seed)
  items = c("low_mood", "fatigue", "worry_a", "worry_b", "poor_sleep", "stress", "loneliness", "inactivity",
    "appetite", "concentration", "support", "optimism")
  x = as.data.table(matrix(rnorm(n * 12), n, 12, dimnames = list(NULL, items)))
  worry = rnorm(n)
  x[, worry_a := as.integer(cut(worry + rnorm(n, sd = 0.3), c(-Inf, -1.2, -0.4, 0.4, 1.2, Inf)))]
  x[, worry_b := pmin(5L, pmax(1L, worry_a + ifelse(runif(n) < 0.02, sample(c(-1L, 1L), n, TRUE), 0L)))]
  eta = -0.3 + 1.2 * x$low_mood + 1.1 * x$fatigue + 0.7 * worry + 0.3 * x$poor_sleep + 0.2 * x$stress +
    0.15 * x$loneliness
  x[, low_wellbeing := factor(rbinom(n, 1, plogis(eta)), levels = c(0, 1))]
  as_task_classif(x, target = "low_wellbeing", positive = "1", id = "survey")
}
task = simulate_survey(500L, seed = 3L)
```

**Fit and look.** XGBoost and ridge logistic regression are fitted on the same five folds as a model without
predictors; held-out PFI is the increase in held-out log loss when one predictor is permuted. Common learners are
labeled by their method in all generated text.

```{r quick-start-fit}
learners = list(
  xgboost = lrn("classif.xgboost", nrounds = 100, eta = 0.1, max_depth = 3, predict_type = "prob"),
  ridge = lrn("classif.cv_glmnet", alpha = 0, nfolds = 5, predict_type = "prob")
)
fits = csdg_fit(task, learners, folds = 5, seed = 1)
imp = csdg_importance(fits, repetitions = 10, seed = 2)
imp
```

```{r quick-start-plot, fig.width = 7, fig.height = 4.5}
plot(imp)
```

**Claim, check, and assess.** The result suggests that XGBoost relies mainly on `low_mood` and `fatigue`.
`claim_relies_mainly()` states the claim, `csdg_check()` computes the properties that it requires, and
`csdg_assess()` applies the decision rule.

```{r quick-start-claim}
claim = claim_relies_mainly(imp, k = 2, factor = 2, learners = "xgboost")
chk = csdg_check(claim, imp)
chk
csdg_assess(chk)
```

The claim was formulated after the explanation was inspected, so its assessment is exploratory; `csdg_confirm(chk,
new_task)` fits the same learners to new data and applies the claim with the same criteria.

## Learn more

* **[Checking what an explanation shows](https://stefancoors.de/mlr3autoiml/articles/exploratory_workflow.html)**
  (`vignette("exploratory_workflow", package = "mlr3autoiml")`) works through the simulated survey with 1,000
  respondents: a claim that is met, a claim about the two wordings that conditional PFI contradicts, a direction,
  and the test on new data.
* **[The numerical example of the article](https://stefancoors.de/mlr3autoiml/articles/claim_scoped_diagnostic_gates.html)**
  (`vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")`) assesses the three claims of the article's
  numerical example.
* **[The function reference](https://stefancoors.de/mlr3autoiml/reference/index.html)** documents every function,
  including the diagnostics of `csdg_audit()` and the local surrogate described in `?csdg_diagnostics`.
* `csdg_claim_record_template()` returns the blank claim record of the article as a table to fill in, and
  `supported()`, `contradicted()`, and `open_status()` return the three property statuses.
* Labels of versions before 0.1.6 still work with a deprecation warning; see
  [the migration note](https://github.com/coorsaa/mlr3autoiml/blob/main/inst/MIGRATION_0_1_6.md) and
  [the changelog](https://stefancoors.de/mlr3autoiml/news/index.html).

## Diagnostics in an mlr3 workflow

`csdg_audit()` derives the gate plan and computes the diagnostics of an mlr3 learner: held-out performance,
dependence and support, calibration, decision curves, cross-fitted local fidelity, stability of held-out PFI,
comparisons with similarly accurate models, leave-one-setting-out refits, and subgroup results.
`csdg_report_card()` lists the gates with the status of each required property, and `csdg_claim_report()` applies
the decision rule to the audit and to evidence records that the researcher supplies.
A numerical criterion needs a source and a rationale; without them, the property remains open.
By default, `csdg_export()` omits fitted models and observation-level predictions; review the generated bundle before
sharing it.

## The accompanying article

Version 0.1.0 produced the primary analyses of the accompanying article. The scripts of its two empirical
applications are kept in a separate analysis repository.

## Legacy AutoIML workflow

The earlier `AutoIML` interface remains available for compatibility. Its gate identifiers and pass/warn/fail/skip
labels are separate from the CSDG gate registry and vocabulary. Its examples and gate overview are in
[the legacy AutoIML article](https://stefancoors.de/mlr3autoiml/articles/legacy_autoiml.html)
(`vignette("legacy_autoiml", package = "mlr3autoiml")`); new analyses start from the CSDG interface above.

## License

LGPL-3.
