---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  fig.retina = 2,
  message   = FALSE,
  warning   = FALSE
)
```

# mlr3autoiml

`mlr3autoiml` is an **auditing and automation layer** for interpretable machine
learning (IML) on top of the **mlr3 ecosystem**. It implements a **gate-based
workflow** (G0A/G0B → G1 → G2 → G3 → G4 → G5 → G6 → G7A/G7B) and assigns
**claim-scoped Interpretation Evidence Levels (IEL-0…3)** that scope the
strength of permissible interpretive claims.

Core dependencies: `mlr3`, `mlr3measures`, `mlr3misc`, `data.table`,
`checkmate`, `R6`. Optional integrations (pipelines, SHAP, iml, plotting)
activate when the corresponding packages are available.

## Installation

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("coorsaa/mlr3autoiml")
```

---

## Example 1 — Classification · decision support

**Task:** `german_credit` (1 000 rows, 20 mixed-type features, binary
`credit_risk`). Full gate path including G3 calibration + decision utility and
G7A subgroup audit.

```{r classif-setup}
library(mlr3)
library(mlr3autoiml)
library(mlr3learners)
library(mlr3pipelines)

task = tsk("german_credit")
task$set_col_roles("personal_status_sex", add_to = "group")

learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 6L)
```

The `ctx$claim` block is a pre-analysis declaration, not a model output.
The operative part is the claim scope, semantics, stakes, and any decision thresholds.
AutoIML can derive conservative defaults for non-use, prohibited interpretations,
and decision-policy wording from those declarations,
so you only need to spell them out when the analysis warrants narrower boundaries.

```{r classif-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "decision_support",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "decision_support",
  semantics = "within_support",
  stakes = "medium",
  claims = list(global = TRUE, local = FALSE, decision = TRUE),
  decision_spec = list(
    thresholds = seq(0.20, 0.60, by = 0.05),
    utility = list(tp = 1, fp = -0.5, fn = -1, tn = 0)
  )
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "personal_status_sex"
auto$ctx$alt_learners = list(
  featureless = lrn("classif.featureless", predict_type = "prob")
)
auto$ctx$multiplicity$transport_mode = "group_performance"
```

The measurement card now only needs the part the analysis cannot infer reliably.
Here that is just the measurement level.
Gate 0B derives task- and pipeline-level notes for missingness and scoring automatically,
and only asks for comparability or reliability evidence when the measurement type makes those claims material.

```{r classif-run, results = "hide"}
res = auto$run()
```

```{r classif-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

```{r classif-iel, echo=FALSE, results='asis'}
cat(sprintf('\n**IEL (overall):** %s\n\n', res$iel$overall))
```

With the operative claim card,
analysis-derived measurement notes, and subgroup audit in place,
the remaining warnings are substantive rather than setup-related:
calibration, net benefit in the declared threshold range,
and the interaction pattern that makes a simple one-feature story too thin.

```{r classif-g3, fig.width = 5.5, fig.height = 4.2, out.width = "49%", fig.show = "hold"}
auto$plot("g3_calibration")
auto$plot("g3_dca")
```

Gate 2 then shows why the package warns instead of presenting a single global
effect curve as the main story:
`amount` and `duration` dominate,
and they interact strongly.

```{r classif-g2, fig.width = 5.5, fig.height = 4.4, out.width = "49%", fig.show = "hold"}
auto$plot("g2_hstats", top_n = 3L)
auto$plot("g2_ale_2d", feature1 = "amount", feature2 = "duration", class_label = task$positive)
```

When Gate 2 regionalization is enabled (`auto$ctx$structure$regionalize = TRUE`) and the interaction screen finds materially heterogeneous effects, the GADGET-style views decompose the global effect into interpretable regions and show the split rules that produced those regions. If no such regions are found, the paper-aligned Gate 2 evidence remains the global effect plot, the interaction screen, and the top 2D ALE surface. Optional `pint_enabled = TRUE` adds a permutation-based interaction screen before regionalization.

```{r classif-g2-gadget, eval = FALSE}
if (!is.null(auto$result$gate_results$G2$artifacts$gadget_regions) &&
    nrow(auto$result$gate_results$G2$artifacts$gadget_regions) > 0L) {
  auto$plot("g2_gadget")
  auto$plot("g2_gadget_tree")
}
auto$plot("g2_pint")
```

```{r classif-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_german_credit", prefix = "german_credit")
```

---

## Example 2 — Regression · global insight

**Task:** `california_housing` (20 433 complete cases from the 20 640-row benchmark,
8 numeric + 1 factor feature, continuous `median_house_value`).
We use a complete-case subset so `ranger`, lasso, and `xgboost` benchmark the same
design matrix. Correlated spatial predictors still trigger ALE selection in G2,
and the mixed learner family makes Gate 6 Rashomon diagnostics realistic.

```{r regr-setup}
task = tsk("california_housing")
task$filter(task$row_ids[stats::complete.cases(task$data())])

learner = lrn(
  "regr.ranger",
  num.trees = 200L,
  mtry.ratio = 0.5,
  min.node.size = 5L
)
```

The claim card here is not extra decoration;
it fixes the semantics and scope before Gate 6 compares models.
Analysis-specific transport and preprocessing notes can remain implicit unless you need
to declare a narrower boundary than the task- and pipeline-level defaults.

```{r regr-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "global_insight",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "global_insight",
  semantics = "within_support",
  stakes = "low",
  claims = list(global = TRUE, local = FALSE, decision = FALSE)
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "ocean_proximity"
auto$ctx$alt_learners = list(
  lasso = as_learner(po("encode") %>>% lrn("regr.glmnet", alpha = 1)),
  xgboost = as_learner(
    po("encode") %>>%
      lrn(
        "regr.xgboost",
        nrounds = 150L,
        eta = 0.05,
        max_depth = 6L,
        subsample = 0.8,
        colsample_bytree = 0.8,
        verbose = 0
      )
  ),
  tree = lrn("regr.rpart", maxdepth = 12L, cp = 1e-03),
  featureless = lrn("regr.featureless")
)
auto$ctx$structure$max_features = 8L
```

```{r regr-run, results = "hide"}
res = auto$run()
```

```{r regr-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

```{r regr-iel, echo=FALSE, results='asis'}
cat(sprintf('\n**IEL (overall):** %s\n\n', res$iel$overall))
```

The effect plots focus on the housing variables that dominate the model story:
income and coastal location.
`ranger` keeps the global effects readable,
while Gate 6 benchmarks it against lasso, `xgboost`, and a single tree baseline.

```{r regr-g2, fig.width = 10, fig.height = 4.4, fig.alt = "ALE effects for median income and spatial coordinates"}
auto$plot("g2_effect", feature = c("median_income", "latitude", "longitude"))
```

The supporting views below add two complementary angles:
conditional SHAP importance for the fitted tree,
and a Rashomon rank heatmap for the near-tie models in Gate 6.

```{r regr-shap-g6, fig.width = 5.8, fig.height = 4.6, out.width = "49%", fig.show = "hold"}
auto$plot("shap_importance", n_rows = 40L, sample_size = 15L, background_n = 40L, top_n = 8L)
auto$plot("g6_rank_heatmap", top_n = 8L)
```

```{r regr-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_california", prefix = "california")
```

---

## Gate overview

| Gate | Triggered when | Evidence checks |
|------|---------------|-----------------|
| G0A  | always | Claim and semantics declaration; hard-stop on `causal_recourse` without identification |
| G0B  | always | Measurement readiness; reliability / invariance evidence for high-stakes |
| G1   | always | CV performance + calibration snapshot |
| G2   | always | Feature dependence · ALE vs PDP selection · pairwise interaction screening |
| G3   | `claims$decision = TRUE` | Calibration curve + Decision Curve Analysis |
| G4   | `claims$local = TRUE` | SHAP faithfulness + perturbation sensitivity checks |
| G5   | always | Stability of narrated patterns under bootstrap resampling |
| G6   | high-resolution profile | Rashomon multiplicity + transport probes |
| G7A  | `sensitive_features` set | Subgroup performance + calibration audit |
| G7B  | user-facing + high-stakes | Human-factors (task-based evaluation) evidence |

IELs (0–3) are assigned per claim scope (global / local / decision).
The overall IEL is the minimum across all requested scopes.

## Status

The package follows a claim-first, gate-based architecture with explicit
semantics, claim-scoped IELs, and evidence-driven gate planning. Default runs
use the `high_resolution` profile; `quick_start = TRUE` is available for rapid
prototyping. IEL-3 requires G6 evidence across all scopes and, for decision
claims, a passed G7A subgroup audit.

## License

LGPL-3.
