---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  fig.retina = 2,
  message   = FALSE,
  warning   = FALSE
)
if (identical(Sys.getenv("CSDG_README_DOCUMENTATION_ONLY"), "true")) {
  knitr::opts_chunk$set(eval = FALSE)
}
```

# mlr3autoiml

`mlr3autoiml` implements **Claim-Scoped Diagnostic Gates (CSDG)** for the **mlr3 ecosystem**. CSDG replaces the
question "What does this explanation show?" with "Does the evidence support this claim?" Its workflow has five steps:
write the claim, the conclusion you intend to report (`csdg_claim()`); specify its scope in six elements, the
quantity, model, procedure, data, meaning, and use; derive the properties the claim requires from the 12 gates
(`csdg_gate_registry()`, `csdg_gate_plan()`); evaluate the evidence on each required property
(`csdg_evidence_record()`, with diagnostics from `csdg_audit()`); and decide and report with a fixed rule
(`csdg_adjudicate_claim()`).

The package records the claim and its scope, derives the gate plan, computes diagnostics, and applies the decision
rule. The researcher sets the criteria, states the relevance of each observation, and enters established
counterevidence and unresolved threats. Favorable results never offset a contradicted property, and no aggregate score
is computed.

| Term | Values |
|---|---|
| Scope elements | quantity, model, procedure, data, meaning (model description, population claim, causal claim), use |
| Evidence roles | required property, established counterevidence, unresolved threat, context |
| Status of a required property | supported, contradicted, open |
| Assessment of a claim | not met if a required property is contradicted; otherwise unresolved if one is open; otherwise met |
| Decision | a met claim is retained; a claim that is not met or unresolved is revised or withheld |

Start with **[the complete walkthrough](vignettes/claim_scoped_diagnostic_gates.Rmd)**
(`vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")`), which follows the five steps with the
numerical example of the accompanying article. Version 0.1.6 uses the vocabulary of the article; labels of earlier
versions still work with a deprecation warning (see [the migration note](inst/MIGRATION_0_1_6.md)).

## Installation

Install the current version from GitHub:

```{r install, eval = FALSE}
remotes::install_github("coorsaa/mlr3autoiml", build_vignettes = TRUE)
library(mlr3autoiml)
vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")
```

Released versions are tagged, for example `remotes::install_github("coorsaa/mlr3autoiml@v0.1.6")`. Version 0.1.0
produced the primary analyses of the accompanying article.

## Quick start: the numerical example of the article

Two items carry the same information: `Z` is standard normal, and both items and the outcome equal `Z`. Model A
predicts with item 1 and model B with item 2 (`f_A(x) = x1`, `f_B(x) = x2`); because the outcome equals each item,
the linear fit in every fold is exact. The held-out PFI with squared error is computed with the package.

```{r quick-start-data}
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3autoiml)
library(data.table)

set.seed(20260926)
n = 10000L
z = rnorm(n)
task = as_task_regr(data.frame(item_1 = z, item_2 = z, y = z), target = "y", id = "numerical_example")
model_using = function(item) {
  as_learner(po("select", selector = selector_name(item), id = paste0("use_", item)) %>>% lrn("regr.lm"))
}
folds = rsmp("cv", folds = 2L)
fits = list(
  A = csdg_resample(task, model_using("item_1"), folds, measures = msrs("regr.mse"), seed = 1L),
  B = csdg_resample(task, model_using("item_2"), folds, measures = msrs("regr.mse"), seed = 1L)
)
groups = list(item_1 = "item_1", item_2 = "item_2", both_items = c("item_1", "item_2"))
pfi = lapply(fits, csdg_fold_pfi, feature_groups = groups, loss = "mse", repetitions = 20L, seed = 2L)
marginal = rbindlist(lapply(names(pfi), function(model) {
  pfi[[model]]$summary[, .(model = model, feature_group, mean_importance)]
}))
dcast(marginal, model ~ feature_group, value.var = "mean_importance")[, lapply(.SD, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})]
```

The exact population values are 2 and 0 in model A and 0 and 2 in model B; permuting both items together (grouped
PFI) gives 2 in both models. Conditional PFI, which draws the permuted item from its distribution given the other
item, is 0 for both items in both models; here the permutation within strata of `Z` implements it, because each item
equals the other.

```{r quick-start-conditional}
conditional = lapply(fits, csdg_fold_pfi, feature_groups = groups[1:2], loss = "mse", repetitions = 5L,
  strata = z, seed = 3L)
rbindlist(lapply(names(conditional), function(model) {
  conditional[[model]]$summary[, .(model = model, feature_group, mean_importance)]
}))
```

**Steps 1 and 2: the claim and its scope.** The claim covers both models.

```{r quick-start-claim}
claim_both = csdg_claim(
  id = "both_models",
  statement = "Under marginal permutation, both models rely more on item 1 than on item 2.",
  claim_type = "global_explanation",
  quantity = "marginal PFI with squared error: increase in expected squared error when one item is permuted",
  model = "several_models",
  procedure = "each item permuted independently of the other item and the outcome; 20 permutations per fold",
  data = "Z standard normal; item 1 = item 2 = outcome = Z; 10,000 simulated observations",
  meaning = "model_description",
  use = "scientific description",
  provenance = list(origin = "specified_before_results", date = "2026-09-26",
    time_basis = "date of the example", selection_basis = "written before the PFI values were computed",
    evidence_ids = character())
)
measurement = csdg_measurement(outcome = "y", predictors = c("item_1", "item_2"), data_source = "simulation",
  sample_definition = "all simulated observations", missingness = "none", preprocessing = "none")
explanation = csdg_explanation(method_ids = "pfi", feature_groups = groups)
```

**Step 3: the required properties.** Besides G0a and G0b, the claim requires G2 (it interprets PFI values), G5 (it
states an ordering), and G6a (it covers two models). Held-out performance (G1) is context, because a model
description holds whatever the model's accuracy.

```{r quick-start-plan}
plan = csdg_gate_plan(claim_both, measurement, explanation)
plan[, .(gate_id, gate_name, required, plan_role)]
```

**Step 4: the evidence.** In model A, the difference between the two items is far beyond the error due to random
permutation (Monte Carlo rule, per fold). The decisive property is G6a: model B reverses the ordering.

```{r quick-start-evidence}
csdg_pfi_mc_difference(pfi$A, "item_1", "item_2")[, .(iteration, estimate, threshold, beyond_monte_carlo_error)]

property = function(gate_id, status, required_property, observation, relevance, criterion = NULL) {
  csdg_evidence_record(
    gate_id, TRUE, "required_property", status = status,
    criterion = if (!is.null(criterion)) list(value = criterion, direction = "qualitative"),
    criterion_source = if (!is.null(criterion)) "translated from the claim",
    criterion_rationale = if (!is.null(criterion)) required_property,
    rationale = observation, required_property = required_property, observation = observation,
    relevance_to_proposition = relevance
  )
}
foundation = list(
  property("G0a", "supported", "The claim and its six scope elements are stated.",
    "All six elements are recorded.", "They fix the quantity, the models, and the procedure."),
  property("G0b", "supported", "The data cover what the claim names.",
    "The population is defined exactly; there is no measurement error or preprocessing.",
    "The claim names no construct and no other population."),
  property("G2", "supported", "Marginal permutation computes the PFI that the claim names.",
    "Squared error and marginal permutation; both models are defined for all inputs.",
    "The claim names the perturbation, so the unrealistic permuted inputs are part of what it describes."),
  property("G5", "supported", "The ordering persists when the quantity is estimated again.",
    "Exact values do not vary.", "The same quantity is computed again.")
)
g6a = property("G6a", "contradicted", "Item 1 has the larger marginal PFI in A and in B.",
  "PFI of item 1 versus item 2: 2 versus 0 in A, 0 versus 2 in B.",
  "Only the model differs between the two computations, and the claim covers both models.",
  criterion = "strict ordering in both models")
```

**Step 5: decide and report.** The original claim is not met and is revised to model A.

```{r quick-start-decision}
assessment_both = csdg_adjudicate_claim(c(foundation, list(g6a)), claim_applicable = TRUE, plan = plan)
assessment_both

claim_a = csdg_claim_revision(
  claim_both, id = "model_a", claim_version = "C1", revision_relation = "narrower",
  statement = "Under marginal permutation, model A relies more on item 1 than on item 2.",
  model = "fitted_model",
  provenance = list(origin = "retrospective_exploratory", date = "2026-09-26", time_basis = "date of the example",
    selection_basis = "restriction chosen after the comparison of A and B", evidence_ids = "both_models_G6a")
)
plan_a = csdg_gate_plan(claim_a, measurement, explanation)
plan_a[required == TRUE, gate_id]
g2_a = property("G2", "supported",
  "Marginal PFI with squared error in model A is larger for item 1, and the wording names the perturbation.",
  "2 versus 0; conditional PFI 0 versus 0.",
  "The computation uses the model, perturbation, and loss that the claim names.",
  criterion = "strict ordering of the exact values; qualifier in the wording")
assessment_a = csdg_adjudicate_claim(c(foundation[c(1L, 2L, 4L)], list(g2_a)), claim_applicable = TRUE,
  plan = plan_a)
assessment_a$assessment
assessment_a$decision_options
```

The revised claim is met (exploratory; it was written after the results were seen) and retained.

## Diagnostics in an mlr3 workflow

`csdg_audit()` derives the gate plan and computes the diagnostics of an mlr3 learner: held-out performance,
dependence and support, calibration, decision curves, cross-fitted local fidelity, stability of held-out PFI,
comparisons with similarly accurate models, leave-one-setting-out refits, and subgroup results.
`csdg_report_card()` lists the gates with the status of each required property (a gate that the claim does not
require is reported as context, without a property status), and `csdg_claim_report()` applies the decision rule to
the audit and to evidence records that the researcher supplies. `csdg_learner_pfi_interval()` computes the corrected
resampled interval of a learner-level mean across folds, and `csdg_pfi_mc_difference()` applies the Monte Carlo rule.
`csdg_local_fidelity_audit()` evaluates local surrogates on points not used to fit them (the fitted surrogate is
described in `?csdg_diagnostics`, section "Local surrogate").
A numerical criterion needs a source and a rationale: for example, `performance = list(maximum_primary_score =
value)` must be accompanied by `criteria = list(performance.maximum_primary_score = list(source = source,
rationale = rationale))`; without them, the property remains open. `csdg_rashomon()` likewise returns candidate
performance without an accepted set unless a tolerance with its source and rationale is supplied.

## Package and study-script boundary

The package owns reusable claim and evidence records, the gate plan, resampling, permutation importance, calibration,
ALE bootstrap, local-fidelity audits, importance agreement, matched-setting utilities, plotting, reporting, and
provenance. Study-specific equal-country sampling, plausible-value loops, SHILD and PISA orchestration, country
exclusion schedules, matched-control scheduling, and manuscript synthesis remain in the analysis scripts, which are
kept in a separate repository. No single package call reproduces either complete empirical study.

Keep `output_dir = NULL` while inspecting an audit, and export to a dedicated directory when it is ready. By default,
`csdg_export()` omits fitted models and observation-level predictions. This is not a privacy or licensing clearance:
review the generated bundle before sharing it, and set `include_models = TRUE` or `include_predictions = TRUE` only
when its release is explicitly approved.

---

## Legacy AutoIML workflow

The examples in this section are not evaluated when this README is rendered. The `AutoIML` interface is the earlier
workflow of the package; its gate identifiers (G0A to G7B, with combined G3 and G6) and its pass/warn/fail/skip labels
are not the CSDG gate registry (`csdg_gate_registry()`) or the CSDG vocabulary.

The earlier `AutoIML` interface remains available for compatibility.
Its `GateResult` pass/warn/fail/skip labels are heuristic workflow diagnostics, not CSDG evidence roles or claim
decisions. Printing a legacy `GateResult` emits a targeted deprecation warning.
The examples below demonstrate that workflow; new analyses should start from
the CSDG interface shown above.

### Example 1 — Classification · decision support

**Task:** `german_credit` (1 000 rows, 20 mixed-type features, binary
`credit_risk`). Full compatibility-gate path including combined G3 calibration and decision utility and
G7A subgroup audit.

```{r classif-setup, eval = FALSE}
library(mlr3)
library(mlr3autoiml)
library(mlr3learners)
library(mlr3pipelines)

task = tsk("german_credit")
task$set_col_roles("personal_status_sex", add_to = "group")

learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 6L)
```

The `ctx$claim` block is a pre-analysis declaration, not a model output.
The operative part is the claim scope, semantics, stakes, and any decision thresholds.
AutoIML can derive conservative defaults for non-use, prohibited interpretations,
and decision-policy wording from those declarations,
so you only need to spell them out when the analysis warrants narrower boundaries.

```{r classif-autoiml, eval = FALSE}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "decision_support",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "decision_support",
  semantics = "within_support",
  stakes = "medium",
  claims = list(global = TRUE, local = FALSE, decision = TRUE),
  decision_spec = list(
    thresholds = seq(0.20, 0.60, by = 0.05),
    utility = list(tp = 1, fp = -0.5, fn = -1, tn = 0)
  )
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "personal_status_sex"
auto$ctx$alt_learners = list(
  featureless = lrn("classif.featureless", predict_type = "prob")
)
auto$ctx$multiplicity$transport_mode = "group_performance"
```

The measurement card now only needs the part the analysis cannot infer reliably.
Here that is just the measurement level.
Gate 0B derives task- and pipeline-level notes for missingness and scoring automatically,
and only asks for comparability or reliability evidence when the measurement type makes those claims material.

```{r classif-run, results = "hide", eval = FALSE}
auto$run()
```

```{r classif-report, eval = FALSE}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

With the operative claim card,
analysis-derived measurement notes, and subgroup audit in place,
the remaining warnings are substantive rather than setup-related:
calibration, net benefit in the declared threshold range,
and the interaction pattern that makes a simple one-feature story too thin.

```{r classif-g3, fig.width = 5.5, fig.height = 4.2, out.width = "49%", fig.show = "hold", eval = FALSE}
auto$plot("g3_calibration")
auto$plot("g3_dca")
```

Gate 2 then shows why the package warns instead of presenting a single global
effect curve as the main story:
`amount` and `duration` dominate,
and they interact strongly.

```{r classif-g2, fig.width = 5.5, fig.height = 4.4, out.width = "49%", fig.show = "hold", eval = FALSE}
auto$plot("g2_hstats", top_n = 3L)
auto$plot("g2_ale_2d", feature1 = "amount", feature2 = "duration", class_label = task$positive)
```

When Gate 2 regionalization is enabled
(`auto$ctx$structure$regionalize = TRUE`) and the interaction screen finds
materially heterogeneous effects, the GADGET-style views decompose the global
effect into interpretable regions and show the split rules that produced those
regions.
If no such regions are found, the paper-aligned Gate 2 evidence remains the
global effect plot, the interaction screen, and the top 2D ALE surface.
Optional `pint_enabled = TRUE` adds a permutation-based interaction screen
before regionalization.

```{r classif-g2-gadget, eval = FALSE}
if (!is.null(auto$result$gate_results$G2$artifacts$gadget_regions) &&
    nrow(auto$result$gate_results$G2$artifacts$gadget_regions) > 0L) {
  auto$plot("g2_gadget")
  auto$plot("g2_gadget_tree")
}
auto$plot("g2_pint")
```

```{r classif-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_german_credit", prefix = "german_credit")
```

---

### Example 2 — Regression · global insight

**Task:** `california_housing` (20 433 complete cases from the 20 640-row benchmark,
8 numeric + 1 factor feature, continuous `median_house_value`).
We use a complete-case subset so `ranger`, lasso, and `xgboost` benchmark the same
design matrix. Correlated spatial predictors still trigger ALE selection in G2,
and the mixed learner family makes Gate 6 Rashomon diagnostics realistic.

```{r regr-setup, eval = FALSE}
task = tsk("california_housing")
task$filter(task$row_ids[stats::complete.cases(task$data())])

learner = lrn(
  "regr.ranger",
  num.trees = 200L,
  mtry.ratio = 0.5,
  min.node.size = 5L
)
```

The claim card here is not extra decoration;
it fixes the semantics and scope before Gate 6 compares models.
Analysis-specific transport and preprocessing notes can remain implicit unless you need
to declare a narrower boundary than the task- and pipeline-level defaults.

```{r regr-autoiml, eval = FALSE}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "global_insight",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "global_insight",
  semantics = "within_support",
  stakes = "low",
  claims = list(global = TRUE, local = FALSE, decision = FALSE)
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "ocean_proximity"
auto$ctx$alt_learners = list(
  lasso = as_learner(po("encode") %>>% lrn("regr.glmnet", alpha = 1)),
  xgboost = as_learner(
    po("encode") %>>%
      lrn(
        "regr.xgboost",
        nrounds = 150L,
        eta = 0.05,
        max_depth = 6L,
        subsample = 0.8,
        colsample_bytree = 0.8,
        verbose = 0
      )
  ),
  tree = lrn("regr.rpart", maxdepth = 12L, cp = 1e-03),
  featureless = lrn("regr.featureless")
)
auto$ctx$structure$max_features = 8L
```

```{r regr-run, results = "hide", eval = FALSE}
auto$run()
```

```{r regr-report, eval = FALSE}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

The effect plots focus on the housing variables that dominate the model story:
income and coastal location.
`ranger` keeps the global effects readable,
while Gate 6 benchmarks it against lasso, `xgboost`, and a single tree baseline.

```{r regr-g2, fig.width = 10, fig.height = 4.4, fig.alt = "ALE effects for median income and spatial coordinates", eval = FALSE}
auto$plot("g2_effect", feature = c("median_income", "latitude", "longitude"))
```

The supporting views below add two complementary angles:
conditional SHAP importance for the fitted tree,
and a Rashomon rank heatmap for the near-tie models in Gate 6.

```{r regr-shap-g6, fig.width = 5.8, fig.height = 4.6, out.width = "49%", fig.show = "hold", eval = FALSE}
auto$plot("shap_importance", n_rows = 40L, sample_size = 15L, background_n = 40L, top_n = 8L)
auto$plot("g6_rank_heatmap", top_n = 8L)
```

```{r regr-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_california", prefix = "california")
```

---

## AutoIML gate overview

This table documents the legacy compatibility interface; it is not the CSDG applicability map returned by
`csdg_gate_plan()` or the gate registry of `csdg_gate_registry()`.

| Gate | Triggered when | Evidence checks |
|------|---------------|-----------------|
| G0A  | always | Claim and semantics declaration; hard-stop on `causal_recourse` without identification |
| G0B  | always | Measurement readiness; reliability / invariance evidence for high-stakes |
| G1   | always | CV performance + calibration snapshot |
| G2   | always | Feature dependence · ALE vs PDP selection · pairwise interaction screening |
| G3   | `claims$decision = TRUE` | Calibration curve + Decision Curve Analysis |
| G4   | `claims$local = TRUE` | SHAP faithfulness + perturbation sensitivity checks |
| G5   | always | Stability of narrated patterns under bootstrap resampling |
| G6   | high-resolution profile | Rashomon multiplicity + transport probes |
| G7A  | `sensitive_features` set | Subgroup performance + calibration audit |
| G7B  | user-facing + high-stakes | Human-factors (task-based evaluation) evidence |

## AutoIML compatibility status

The compatibility workflow uses a claim-first, gate-based architecture with
explicit semantics, direct gate evidence, and claim-dependent gate planning.
Default compatibility runs use the `high_resolution` profile;
`quick_start = TRUE` is available for rapid prototyping. Report cards and
guides retain the individual gate results and their claim-specific restrictions
instead of reducing them to a summary evidence score.

## License

LGPL-3.
