---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  fig.retina = 2,
  message   = FALSE,
  warning   = FALSE
)
if (identical(Sys.getenv("CSDG_README_DOCUMENTATION_ONLY"), "true")) {
  knitr::opts_chunk$set(eval = FALSE)
}
```

# mlr3autoiml

Start with **[the complete claim-to-evidence walkthrough](vignettes/claim_scoped_diagnostic_gates.Rmd)**.
After installation (see below), open `vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")`.
It works through the numerical example of the accompanying article (its two claims plus an unresolved third claim),
a small mlr3 audit, and an export and read-back check.

`mlr3autoiml` 0.1.5 primarily implements **Claim-Scoped Diagnostic Gates
(CSDG)** for interpretable machine-learning analyses in the **mlr3 ecosystem**.
A CSDG audit records the intended claim, measurement and preprocessing choices,
explanation semantics, gate-specific evidence, and the limits of the resulting
interpretation. Module availability, evidence role, result direction, criterion provenance, materiality, and claim
consequence are separate fields. CSDG does not calculate an aggregate score, readiness grade, or Interpretation
Evidence Level, and it does not certify causal validity, fairness, or deployment readiness.
Claims use exactly three inference levels (`functional`, `predictive`, and `substantive`). A separate `use_claim`
flag records whether the proposition additionally asserts adequacy for an audience, workflow, implementation, or use.

The claim-scoped interface distinguishes calibration from utility (G3a/G3b), model multiplicity from setting
transport (G6a/G6b), and technical subgroup behavior from audience and workflow evidence (G7a/G7b). The earlier
combined gate names below belong only to the retained `AutoIML` compatibility interface.

The claim object is `C = (phi, kappa)`: `statement` records the proposition, while `kappa = (T, M, S, D, U, Q)` records
target, model scope, explanation semantics, analytic distribution and measurement context, intended use, and explanation
design. Matching contexts cannot establish matching propositions. `csdg_claim_relation()` separates the declared
context comparison from an explicit proposition comparison; without the latter, `relation` is `unchecked`.
Neither a declaration nor a complete record is a logical proof. Evidence roles
are exactly `necessary_requirement`, `potential_defeater`, `graded_support`, and `descriptive_context`.
`csdg_evidence_record()` keeps those roles distinct from completion and results, and `csdg_adjudicate_claim()` applies
conditional non-compensation without adding evidence into a score.

Claim applicability is declared independently of the supplied modules. An in-scope claim with no relevant evidence
is unresolved, not out of scope. Recorded consequences constrain adjudication; unresolved evidence cannot silently
become a favorable decision. See [the 0.1.4 contract migration note](inst/MIGRATION_0_1_4.md).

Sensitivity records distinguish the varied component, what stays fixed, whether the estimand is unchanged, and whether
the proposition claims invariance. A legitimate change in question is not automatically counterevidence. New
claim-constraining variation requires a required-property, observation, and proposition-relevance chain; changed or
uncertain estimands additionally require explicitly claimed invariance. Legacy records remain readable with linkage
marked `not_recorded`, not retrospectively validated. Claim provenance can record prior specification, exploratory
revision, or independently confirmed evidence; these are documentary origins, not interpretation levels.

`csdg_resolve_sources()` verifies case-exact public paths and declared protected or external references without opening
protected records or fetching external resources. Availability and checksum agreement do not prove evidentiary support.
See [migration notes](inst/MIGRATION_0_1_2.md) for interface examples and the explicit context-only compatibility change.

Core dependencies: `mlr3`, `mlr3measures`, `mlr3misc`, `data.table`,
`checkmate`, `R6`. Optional integrations (pipelines, SHAP, iml, plotting)
activate when the corresponding packages are available.

## Installation

Install the current version from GitHub:

```{r install, eval = FALSE}
remotes::install_github("coorsaa/mlr3autoiml", build_vignettes = TRUE)
library(mlr3autoiml)
vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")
```

Released versions are tagged. Version 0.1.0 produced the primary analyses of the accompanying article; install a
tagged version with, for example, `remotes::install_github("coorsaa/mlr3autoiml@v0.1.5")`.

## CSDG quick start

Declare the question, estimand, and scope before evaluating the diagnostic results, rather than prescribing a desired
positive finding. Exploratory revisions are legitimate when their selection basis remains explicit.
This example requests descriptive held-out performance, calibration, and global permutation-importance evidence for
one selected model in the analytic sample. Because it supplies no use-linked adequacy criterion, the corresponding
adequacy questions remain unresolved.

Prediction-based triage requires appropriate prediction, calibration, use, and consequence evidence, not automatically
causal effects of every predictor. Advice to change a predictor to improve an outcome requires an intervention argument.

```{r csdg-quick-start, eval = FALSE}
library(mlr3)
library(mlr3autoiml)
library(mlr3learners)

task = tsk("german_credit")
learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 6L)

claim = csdg_claim(
  id = "credit_risk_description",
  statement = paste(
    "Held-out predictive performance, calibration, and global permutation importance",
    "are described for the selected model in the analytic sample."
  ),
  claim_type = c("predictive_performance", "calibration", "global_explanation"),
  target = "credit risk",
  unit = "credit application",
  population = "applications represented by the analytic sample",
  analytic_distribution = "the observed German credit example data",
  model_scope = "selected_model",
  setting_scope = "analytic_sample",
  scientific_use = "descriptive model audit",
  explanation_design = "held-out marginal permutation importance"
)

measurement = csdg_measurement(
  outcome = task$target_names,
  predictors = task$feature_names,
  data_source = "mlr3 German credit example task",
  sample_definition = "all rows in the example task",
  unit = "credit application",
  missingness = "no missing values in the supplied task",
  preprocessing = "the complete learner is refitted within every training split",
  verification = list(status = "not_checked")
)

explanation = csdg_explanation(
  method_ids = "pfi",
  scope = "global",
  target = "positive-class probability"
)

audit = csdg_audit(
  task = task,
  learner = learner,
  claim = claim,
  measurement = measurement,
  explanation = explanation,
  config = csdg_config(
    seed = 42L,
    resampling = list(folds = 3L, repeats = 1L),
    stability = list(pfi_repetitions = 3L)
  )
)

csdg_report_card(audit)
csdg_claim_report(audit)
bundle = csdg_export(audit, path = "csdg-output")
```

When a numerical criterion is appropriate, supply both the value and its provenance. For example,
`performance = list(maximum_primary_score = value)` must be accompanied by
`criteria = list(performance.maximum_primary_score = list(source = source, rationale = rationale))`.
Without both strings, the diagnostic remains descriptive and the module decision is `unresolved`.

`csdg_rashomon()` is likewise descriptive by default. It returns candidate performance without an accepted set unless
an absolute or relative tolerance and its source and rationale are supplied explicitly.

## Package and study-script boundary

The package owns reusable claim and evidence records, gate planning, resampling, permutation importance, calibration,
ALE bootstrap, local-fidelity audits, importance agreement, matched-setting utilities, plotting, reporting, and
provenance. Study-specific equal-country sampling, plausible-value loops, SHILD and PISA orchestration, country
exclusion schedules, matched-control scheduling, and manuscript synthesis remain in the analysis scripts, which are kept in a separate repository. No single
package call reproduces either complete empirical study.

Keep `output_dir = NULL` while inspecting an audit, and export to a dedicated
directory when it is ready. By default, `csdg_export()` omits fitted models and
observation-level predictions. This is not a privacy or licensing clearance:
review the generated bundle before sharing it, and set `include_models = TRUE`
or `include_predictions = TRUE` only when its release is explicitly approved.

---

## AutoIML compatibility examples

The earlier `AutoIML` interface remains available for compatibility.
Its `GateResult` pass/warn/fail/skip labels are heuristic workflow diagnostics, not CSDG evidence roles or claim
decisions. Printing a legacy `GateResult` emits a targeted deprecation warning.
The examples below demonstrate that workflow; new analyses should start from
the CSDG claim-card and audit interface shown above.

### Example 1 — Classification · decision support

**Task:** `german_credit` (1 000 rows, 20 mixed-type features, binary
`credit_risk`). Full compatibility-gate path including combined G3 calibration and decision utility and
G7A subgroup audit.

```{r classif-setup}
library(mlr3)
library(mlr3autoiml)
library(mlr3learners)
library(mlr3pipelines)

task = tsk("german_credit")
task$set_col_roles("personal_status_sex", add_to = "group")

learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 6L)
```

The `ctx$claim` block is a pre-analysis declaration, not a model output.
The operative part is the claim scope, semantics, stakes, and any decision thresholds.
AutoIML can derive conservative defaults for non-use, prohibited interpretations,
and decision-policy wording from those declarations,
so you only need to spell them out when the analysis warrants narrower boundaries.

```{r classif-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "decision_support",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "decision_support",
  semantics = "within_support",
  stakes = "medium",
  claims = list(global = TRUE, local = FALSE, decision = TRUE),
  decision_spec = list(
    thresholds = seq(0.20, 0.60, by = 0.05),
    utility = list(tp = 1, fp = -0.5, fn = -1, tn = 0)
  )
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "personal_status_sex"
auto$ctx$alt_learners = list(
  featureless = lrn("classif.featureless", predict_type = "prob")
)
auto$ctx$multiplicity$transport_mode = "group_performance"
```

The measurement card now only needs the part the analysis cannot infer reliably.
Here that is just the measurement level.
Gate 0B derives task- and pipeline-level notes for missingness and scoring automatically,
and only asks for comparability or reliability evidence when the measurement type makes those claims material.

```{r classif-run, results = "hide"}
auto$run()
```

```{r classif-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

With the operative claim card,
analysis-derived measurement notes, and subgroup audit in place,
the remaining warnings are substantive rather than setup-related:
calibration, net benefit in the declared threshold range,
and the interaction pattern that makes a simple one-feature story too thin.

```{r classif-g3, fig.width = 5.5, fig.height = 4.2, out.width = "49%", fig.show = "hold"}
auto$plot("g3_calibration")
auto$plot("g3_dca")
```

Gate 2 then shows why the package warns instead of presenting a single global
effect curve as the main story:
`amount` and `duration` dominate,
and they interact strongly.

```{r classif-g2, fig.width = 5.5, fig.height = 4.4, out.width = "49%", fig.show = "hold"}
auto$plot("g2_hstats", top_n = 3L)
auto$plot("g2_ale_2d", feature1 = "amount", feature2 = "duration", class_label = task$positive)
```

When Gate 2 regionalization is enabled
(`auto$ctx$structure$regionalize = TRUE`) and the interaction screen finds
materially heterogeneous effects, the GADGET-style views decompose the global
effect into interpretable regions and show the split rules that produced those
regions.
If no such regions are found, the paper-aligned Gate 2 evidence remains the
global effect plot, the interaction screen, and the top 2D ALE surface.
Optional `pint_enabled = TRUE` adds a permutation-based interaction screen
before regionalization.

```{r classif-g2-gadget, eval = FALSE}
if (!is.null(auto$result$gate_results$G2$artifacts$gadget_regions) &&
    nrow(auto$result$gate_results$G2$artifacts$gadget_regions) > 0L) {
  auto$plot("g2_gadget")
  auto$plot("g2_gadget_tree")
}
auto$plot("g2_pint")
```

```{r classif-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_german_credit", prefix = "german_credit")
```

---

### Example 2 — Regression · global insight

**Task:** `california_housing` (20 433 complete cases from the 20 640-row benchmark,
8 numeric + 1 factor feature, continuous `median_house_value`).
We use a complete-case subset so `ranger`, lasso, and `xgboost` benchmark the same
design matrix. Correlated spatial predictors still trigger ALE selection in G2,
and the mixed learner family makes Gate 6 Rashomon diagnostics realistic.

```{r regr-setup}
task = tsk("california_housing")
task$filter(task$row_ids[stats::complete.cases(task$data())])

learner = lrn(
  "regr.ranger",
  num.trees = 200L,
  mtry.ratio = 0.5,
  min.node.size = 5L
)
```

The claim card here is not extra decoration;
it fixes the semantics and scope before Gate 6 compares models.
Analysis-specific transport and preprocessing notes can remain implicit unless you need
to declare a narrower boundary than the task- and pipeline-level defaults.

```{r regr-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "global_insight",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "global_insight",
  semantics = "within_support",
  stakes = "low",
  claims = list(global = TRUE, local = FALSE, decision = FALSE)
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "ocean_proximity"
auto$ctx$alt_learners = list(
  lasso = as_learner(po("encode") %>>% lrn("regr.glmnet", alpha = 1)),
  xgboost = as_learner(
    po("encode") %>>%
      lrn(
        "regr.xgboost",
        nrounds = 150L,
        eta = 0.05,
        max_depth = 6L,
        subsample = 0.8,
        colsample_bytree = 0.8,
        verbose = 0
      )
  ),
  tree = lrn("regr.rpart", maxdepth = 12L, cp = 1e-03),
  featureless = lrn("regr.featureless")
)
auto$ctx$structure$max_features = 8L
```

```{r regr-run, results = "hide"}
auto$run()
```

```{r regr-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

The effect plots focus on the housing variables that dominate the model story:
income and coastal location.
`ranger` keeps the global effects readable,
while Gate 6 benchmarks it against lasso, `xgboost`, and a single tree baseline.

```{r regr-g2, fig.width = 10, fig.height = 4.4, fig.alt = "ALE effects for median income and spatial coordinates"}
auto$plot("g2_effect", feature = c("median_income", "latitude", "longitude"))
```

The supporting views below add two complementary angles:
conditional SHAP importance for the fitted tree,
and a Rashomon rank heatmap for the near-tie models in Gate 6.

```{r regr-shap-g6, fig.width = 5.8, fig.height = 4.6, out.width = "49%", fig.show = "hold"}
auto$plot("shap_importance", n_rows = 40L, sample_size = 15L, background_n = 40L, top_n = 8L)
auto$plot("g6_rank_heatmap", top_n = 8L)
```

```{r regr-export, eval = FALSE}
export_analysis_bundle(auto, dir = "bundle_california", prefix = "california")
```

---

## AutoIML gate overview

This table documents the legacy compatibility interface; it is not the CSDG applicability map returned by
`csdg_gate_plan()`.

| Gate | Triggered when | Evidence checks |
|------|---------------|-----------------|
| G0A  | always | Claim and semantics declaration; hard-stop on `causal_recourse` without identification |
| G0B  | always | Measurement readiness; reliability / invariance evidence for high-stakes |
| G1   | always | CV performance + calibration snapshot |
| G2   | always | Feature dependence · ALE vs PDP selection · pairwise interaction screening |
| G3   | `claims$decision = TRUE` | Calibration curve + Decision Curve Analysis |
| G4   | `claims$local = TRUE` | SHAP faithfulness + perturbation sensitivity checks |
| G5   | always | Stability of narrated patterns under bootstrap resampling |
| G6   | high-resolution profile | Rashomon multiplicity + transport probes |
| G7A  | `sensitive_features` set | Subgroup performance + calibration audit |
| G7B  | user-facing + high-stakes | Human-factors (task-based evaluation) evidence |

## AutoIML compatibility status

The compatibility workflow uses a claim-first, gate-based architecture with
explicit semantics, direct gate evidence, and claim-dependent gate planning.
Default compatibility runs use the `high_resolution` profile;
`quick_start = TRUE` is available for rapid prototyping. Report cards and
guides retain the individual gate results and their claim-specific restrictions
instead of reducing them to a summary evidence score.

## License

LGPL-3.
