---
title: "The numerical example of the article"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The numerical example of the article}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
run = all(vapply(c("mlr3learners", "mlr3pipelines"), requireNamespace, logical(1L), quietly = TRUE))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = run)
txt = function(x) if (run) x else "?"
```

The article illustrates Claim-Scoped Diagnostic Gates (CSDG) with two items that carry the same information. Let `Z`
be standard normal, let both items and the outcome equal `Z`, and use squared error. Model A predicts with item 1,
`f_A(x) = x1`, and model B with item 2, `f_B(x) = x2`. In model A, shuffling item 1 replaces the prediction `Z` with
an independent draw `Z'`, so the expected squared error rises from 0 to `E[(Z - Z')^2] = 2`; shuffling item 2 leaves
the prediction unchanged, so its permutation feature importance (PFI) is 0. Model B reverses the values. Grouped
PFI, which shuffles both items together, is 2 in both models; conditional PFI, which draws the permuted item given
the other item, is 0, because that draw equals the original value. This vignette computes these values from 10,000
simulated observations and assesses the three claims of the article with the functions introduced in
`vignette("exploratory_workflow", package = "mlr3autoiml")`.
`r if (!run) "The code is not run because the suggested packages mlr3learners or mlr3pipelines are not installed."`

```{r packages, message = FALSE, warning = FALSE}
library(mlr3autoiml)
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
```

## Data, models, and PFI

Each model is a linear model of one item, which reproduces the outcome exactly in every fold. The supplied strata
are the exact values of `Z` (`strata`), so each stratum is one row: permuting an item within its stratum leaves it
unchanged, which is the conditional distribution of one item given the other, and conditional PFI is exactly 0. The
package warns about such small strata, because in real data they usually mean that the permutation changes almost
nothing.

```{r importance}
set.seed(1)
z = rnorm(10000)
task = as_task_regr(data.frame(item_1 = z, item_2 = z, y = z), target = "y", id = "numerical_example")
model_using = function(item) {
  as_learner(po("select", selector = selector_name(item), id = paste0("use_", item)) %>>% lrn("regr.lm"))
}
fits = csdg_fit(task, list(A = model_using("item_1"), B = model_using("item_2")), folds = 5,
  measure = "regr.mse", labels = c(A = "model A", B = "model B"), seed = 1)
imp = csdg_importance(fits, repetitions = 20, groups = list(both_items = c("item_1", "item_2")),
  strata = list(item_1 = z, item_2 = z), seed = 2)
imp
```

The estimates are close to the exact values; relative to the improvement over the model without predictors, they
are about 200%, because PFI values are not parts of the improvement. The claims below are derived from these results, so
the package labels their assessments exploratory.

## The original claim covers both models

```{r claim-both}
items = c(item_1 = "item 1", item_2 = "item 2")
claim_both = claim_order(imp, "item_1", "item_2", learners = c("A", "B"), marginal_only = TRUE, labels = items)
res_both = csdg_assess(csdg_check(claim_both, imp))
res_both
```

```{r assert-both, include = FALSE}
stopifnot(identical(res_both$assessment, "not_met"), "G6a" %in% res_both$decisive$gate)
```

The claim is `r txt(res_both$label)`: it covers both models, so its content is checked for each under G6a, and
model B reverses the ordering.

## The qualifier is needed

On realistic inputs, where the items are equal, A and B are the same function; only a permuted input has unequal
items. Without the qualifier "under marginal permutation," the claim about model A states more than marginal PFI
shows.

```{r claim-a}
claim_a = claim_order(imp, "item_1", "item_2", learners = "A", labels = items)
res_a = csdg_assess(csdg_check(claim_a, imp))
res_a
```

```{r assert-a, include = FALSE}
stopifnot(identical(res_a$assessment, "not_met"), "G2" %in% res_a$decisive$gate)
```

Conditional PFI is 0 for both items, so the procedure property (`r txt(unique(res_a$decisive$gate))`) is
contradicted and the claim is `r txt(res_a$label)`.

## The revised claim

The revised claim names model A and the marginal permutation.

```{r claim-revised}
claim_revised = claim_order(imp, "item_1", "item_2", learners = "A", marginal_only = TRUE, labels = items)
chk_revised = csdg_check(claim_revised, imp)
chk_revised
res_revised = csdg_assess(chk_revised)
res_revised
```

```{r assert-revised, include = FALSE}
stopifnot(identical(res_revised$assessment, "met"))
```

The revised claim is `r txt(res_revised$label)`. Model B's reversal is context, because the claim no longer covers
model B, and the grouped PFI measures reliance on the pair rather than ordering its members.

## Further reading

In the article's variant with items of different reliability, the model that uses the more reliable item has the
larger PFI, although both items measure the same `Z`; importance can then reflect the reliability of a measure
rather than the relevance of its construct. Analysts who record claims, scopes, and evidence by hand, for example
for a claim stated before the analysis, find the record functions in `?csdg_claim`, `?csdg_evidence_record`,
`?csdg_adjudicate_claim`, and `?csdg_claim_relation`.
