---
title: "From a claim to evidence: a complete CSDG walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{From a claim to evidence: a complete CSDG walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = TRUE)
```

This guide implements the five steps of the accompanying paper:
**write the claim, specify the scope, derive the required properties, evaluate the evidence, and decide and report**.
The input is a conclusion a researcher wants to report, the claim.
The output is the assessment of that claim (met, not met, or unresolved), the argument behind it, and what can be
reported. The researcher chooses the required properties; the package records them and applies the
decision rule. It does not discover all relevant properties or judge their scientific adequacy.

All code below runs on generated data. No SHILD or PISA files are needed.
The first example is the numerical example of Section 2.6 of the paper.
A short fitted-model example then connects the record to `mlr3`, followed by an export and read-back check.

```{r package, warning = FALSE}
library(mlr3autoiml)
library(data.table)
library(mlr3)
library(mlr3learners)
packageVersion("mlr3autoiml")
```

<style>
pre, pre code, pre > code.sourceCode { white-space: pre-wrap; overflow-wrap: anywhere; }
pre > code.sourceCode { display: block; }
pre > code.sourceCode > span { display: inline-block; max-width: 100%; white-space: pre-wrap; overflow-wrap: anywhere; }
table { width: 100%; table-layout: auto; }
th, td { overflow-wrap: anywhere; vertical-align: top; }
</style>

## 1. Write the claim

Two predictors carry identical information: let `Z` be standard normal and set `X1 = X2 = Y = Z`.
Model A returns predictor 1 and model B returns predictor 2, for every input; both predict the outcome without error.
The claim is: **both models rely more on predictor 1 than on predictor 2**. Its scope measures reliance by marginal
permutation feature importance (PFI) with squared error. The claim describes two explicitly defined prediction
functions, not the causal importance of either predictor. The record also notes that the claim was written before the values
were computed.

```{r claim}
claim_both = csdg_claim(
  id = "both_models",
  statement = "Both models rely more on predictor 1 than on predictor 2.",
  claim_type = c("global_explanation", "model_generalization"),
  target = "increase in mean squared prediction error under marginal permutation",
  semantics = "hypothetical_model_query",
  unit = "generated observation",
  population = "Z standard normal; X1 = X2 = Y = Z",
  analytic_distribution = "the specified standard-normal population",
  model_scope = "near_equivalent_models",
  scientific_use = "describe the two specified prediction functions",
  explanation_design = "compare fA(x) = x1 and fB(x) = x2 under the same marginal permutation",
  claim_level = "functional",
  notes = "Two constructed equally accurate functions; no empirical model-selection claim.",
  provenance = list(origin = "specified_before_results", date = "2026-09-26",
    time_basis = "tutorial construction date", selection_basis = "written before the PFI values were computed",
    evidence_ids = character())
)
claim_both$statement
```

## 2. Specify the scope

The call above records the six scope elements of the paper: the quantity (`target`), the model (`model_scope`),
the procedure (`explanation_design`), the data (`analytic_distribution`), the meaning (`semantics`), and the use
(`scientific_use`). The meaning `hypothetical_model_query` marks a model description under perturbed inputs:
marginal permutation evaluates the models at inputs where the two predictors differ, which never occur in the data.
The measurement and explanation records add details of the data and the procedure.

```{r scope}
measurement = csdg_measurement(
  outcome = "y", predictors = c("x1", "x2"),
  data_source = "specified synthetic distribution",
  sample_definition = "all generated observations; no exclusions",
  unit = "generated observation", outcome_scale = "continuous, variance one",
  missingness = "none", preprocessing = "none"
)
explanation = csdg_explanation(
  method_ids = "pfi", scope = "global", target = "mean squared prediction error",
  perturbation = list(type = "marginal", distribution = "independent standard-normal copy")
)
```

## 3. Derive the required properties

A gate is a generic evidence question; applied to a claim, it becomes a required property.
Following Table 3 of the paper, the claim requires the specification and measurement gates (G0a, G0b), the procedure
gate (G2) because it interprets PFI values, the stability gate (G5) because it compares them, and the models gate
(G6a) because it covers two models. Calibration and local fidelity are not required.

The package suggests a plan for the declared claim type. The plan is a starting point, not a verdict, and its default
roles are weaker than the paper's: it proposes G2 as possible counterevidence and G5 as context, and it adds
predictive performance (G1). We therefore align the plan with the paper before recording evidence.

```{r selection}
plan = csdg_gate_plan(claim_both, measurement, explanation)
plan[gate_id == "G1", `:=`(applicable = FALSE, evidence_role = "descriptive_context")]
plan[gate_id %in% c("G2", "G5"), evidence_role := "necessary_requirement"]
knitr::kable(plan[applicable == TRUE, c("gate_id", "gate_name", "evidence_role"), with = FALSE])
```

G1 is not required because the claim describes two functions that predict without error by construction; it is not
a population claim and does not use predicted values. The software labels of some gates differ from the names
in the paper (for example, G6a is labeled "Model multiplicity" here and "Models" in the paper).

## 4. Evaluate the evidence

For each required property, the record states the property, the observation, and its relevance: why the observation
bears on the claim; decisive properties also record their criterion. Shuffling the used predictor replaces the prediction `Z` with an independent draw `Znew`, so the
expected squared error rises from 0 to `E[(Z - Znew)^2] = Var(Z) + Var(Znew) = 2`, because the two are independent
with mean 0 and variance 1; shuffling the unused predictor changes nothing.
These are exact population values. The following Monte Carlo calculation illustrates them with both functions fixed.

```{r compute}
set.seed(260926)
n = 10000L
z = rnorm(n)
znew = rnorm(n)
f_a = function(x1, x2) x1
f_b = function(x1, x2) x2
mse = function(prediction) mean((z - prediction)^2)
observed = data.frame(
  model = c("A", "B"),
  predictor_1 = c(mse(f_a(znew, z)), mse(f_b(znew, z))),
  predictor_2 = c(mse(f_a(z, znew)), mse(f_b(z, znew))))
knitr::kable(observed, digits = 4,
  caption = "Finite Monte Carlo estimates; the exact population rows are (2, 0) and (0, 2).")
stopifnot(observed$predictor_1[1L] > observed$predictor_2[1L],
  observed$predictor_1[2L] < observed$predictor_2[2L])
```

The foundation, procedure, and stability properties are supported: the scope is complete, the population is defined
exactly, marginal permutation computes the named quantity, and exact values cannot change when recomputed.

```{r evidence-supported}
supported = function(gate_id, property, observation, relevance, criterion = NULL, criterion_rationale = NULL) {
  csdg_evidence_record(
    gate_id, TRUE, "necessary_requirement", result_direction = "supports",
    claim_consequence = "retain_exact_claim", rationale = observation,
    criterion = criterion, criterion_source = if (!is.null(criterion)) "translated from the claim",
    criterion_rationale = criterion_rationale,
    required_property = property, observation = observation, relevance_to_proposition = relevance
  )
}
foundation = list(
  supported("G0a", "The claim and all six scope elements are stated.",
    "csdg_claim() records all six elements.", "They fix the quantity, the models, and the procedure."),
  supported("G0b", "The data match what the claim says.",
    "The population is defined exactly; there is no measurement error or preprocessing.",
    "The claim names no construct and no other population."),
  supported("G2", "Marginal permutation computes the PFI named in the claim.",
    "Squared-error loss and marginal permutation; both functions are defined for all inputs.",
    "The claim describes A and B as functions, so the unrealistic permuted inputs are part of what it describes."),
  supported("G5", "The ordering persists when the quantity is computed again.",
    "The values are exact population values.", "Exact values cannot change when recomputed.")
)
```

The decisive property is contradicted: only the model differs between the two computations, the claim covers both
models, and model B reverses the ordering. The software records this as `challenges`. The label records the
researcher's argument; it is not inferred by the software from the numbers.

```{r evidence-both}
evidence_g6a = csdg_evidence_record(
  "G6a", TRUE, "necessary_requirement", result_direction = "challenges",
  claim_consequence = "revise_claim",
  criterion = list(value = "strict ordering in both models", direction = "qualitative"),
  criterion_source = "translated from the claim",
  criterion_rationale = "The claim asserts the ordering in each of the two models.",
  rationale = "The exact values are (2, 0) in A and (0, 2) in B; the stated ordering fails in B.",
  varied_component = "prediction function: A versus B",
  held_constant = c("population", "squared-error loss", "marginal permutation"),
  same_estimand = FALSE,
  same_estimand_rationale = "Each function has its own PFI; the claim explicitly covers both.",
  invariance_claimed = TRUE,
  required_property = "Predictor 1 has the larger marginal PFI in each of the two models.",
  observation = "Exact population values: A = (2, 0); B = (0, 2).",
  relevance_to_proposition = "Only the model differs, and the claim covers model B, which reverses the ordering."
)
evidence_both = c(foundation, list(evidence_g6a))
decision_both = csdg_adjudicate_claim(evidence_both, claim_applicable = TRUE)
decision_both$decision
stopifnot(identical(decision_both$decision, "not_met"))
```

Other versions of PFI answer other questions. Conditional PFI replaces predictor 1 with a draw from its distribution
given predictor 2; because the predictors are identical, the draw equals the original value, so conditional PFI is 0.
Grouped PFI shuffles both predictors together and gives 2 for either function.
These values do not contradict the marginal values; they answer different questions.

## 5. Decide and report

The original claim is not met, so it is revised. The revised claim, **model A relies more on predictor 1 than on
predictor 2**, is a new claim with its own record and assessment. It follows logically from the original claim, but
because the original claim is not met, it needs its own evidence, which the values for model A supply. The record notes that the revision was chosen
after the original comparison had been examined.

```{r revised-claim}
claim_a = csdg_claim_revision(
  claim_both, id = "model_a",
  statement = "Model A relies more on predictor 1 than on predictor 2.",
  claim_version = "revised", revision_relation = "narrower",
  claim_type = "global_explanation", model_scope = "selected_model",
  explanation_design = "fA(x) = x1; marginal permutation; squared-error loss",
  provenance = list(origin = "retrospective_exploratory", date = "2026-09-26",
    time_basis = "tutorial construction date",
    selection_basis = "restriction motivated by the two-model comparison", evidence_ids = "both_models_G6a")
)
evidence_a = c(foundation[c(1L, 2L, 4L)], list(supported("G2",
  "The computation is marginal PFI with squared error in model A, and it yields the larger value for predictor 1.",
  "Exact values in A: 2 versus 0.",
  paste("The computation uses the model, perturbation, and loss named in the claim;",
    "the claim describes A as a function, so the unrealistic permuted inputs are part of what it describes."),
  criterion = list(value = "strict ordering of the exact values", direction = "qualitative"),
  criterion_rationale = "The claim asserts that model A relies more on predictor 1 than on predictor 2.")))
decision_a = csdg_adjudicate_claim(evidence_a, claim_applicable = TRUE)
decision_a$decision
stopifnot(identical(decision_a$decision, "met"))
relation = csdg_claim_relation(
  claim_both, claim_a,
  coordinate_relations = c(target = "same", model_scope = "narrower", semantics = "same",
    analytic_distribution = "same", scientific_use = "same", explanation_design = "narrower"),
  rationale = "The claim about A and B is restricted to A under the same loss and distribution.",
  proposition_relation = "logical_weakening",
  proposition_rationale = "If the ordering holds in both A and B, it holds in A. This implication does not assert the premise.",
  revision_kind = "logical_weakening"
)
relation$proposition_relation
```

The package stores this reasoning; it does not prove the implication.

The vignette adds a third claim that the paper does not discuss, to show an unresolved assessment:
**the explanations can guide decisions about individual cases**. It interprets the PFI explanation (G2) and requires evidence
that acting on the explanation improves such decisions (G3b), judged against a criterion derived from the decision.
Neither the decision nor the criterion is specified, so its scope is incomplete (G0a open), and G3b is open. G2 is
open as well: without a specified decision, it is unknown which quantity the decision needs and whether it concerns
only realistic inputs, where A and B coincide. All three gaps are missing evidence, not contrary evidence, so the claim
is unresolved.

```{r unresolved-use}
claim_use = csdg_claim(
  id = "individual_decisions",
  statement = "The explanations can guide decisions about individual cases.",
  claim_type = "decision", use_claim = TRUE,
  target = "adequacy for an individual decision", claim_level = "functional",
  scientific_use = "individual use has not yet been specified"
)
open_g0a = csdg_evidence_record(
  "G0a", TRUE, "necessary_requirement", availability = "incomplete",
  result_direction = "not_evaluated", claim_consequence = "unresolved",
  rationale = "The decision and its criterion are not specified.",
  required_property = "The claim and all six scope elements are stated.",
  observation = "The use element names no decision.",
  relevance_to_proposition = "Without the use, the required decision evidence cannot be defined."
)
open_g2 = csdg_evidence_record(
  "G2", TRUE, "necessary_requirement", availability = "incomplete",
  result_direction = "not_evaluated", claim_consequence = "unresolved",
  rationale = "The use is not specified.",
  required_property = "The procedure computes the quantity that the decision needs.",
  observation = "Marginal permutation evaluates the models at inputs where x1 and x2 differ.",
  relevance_to_proposition = "Whether these unrealistic inputs matter depends on the decision, which is not specified."
)
evidence_use = c(list(open_g0a, open_g2), foundation[2L], list(csdg_evidence_record(
  "G3b", TRUE, "necessary_requirement", availability = "unavailable",
  result_direction = "not_evaluated", claim_consequence = "unresolved",
  rationale = "No criterion derived from a decision and no evidence on decisions are available.",
  required_property = "Acting on the explanation improves decisions about individuals.",
  observation = "No criterion derived from a decision and no evidence on decisions.",
  relevance_to_proposition = "PFI values describe the models, not the consequences of decisions."
)))
decision_use = csdg_adjudicate_claim(evidence_use, claim_applicable = TRUE)
stopifnot(identical(decision_use$decision, "unresolved"))
reported_claims = c(
  "Not retained; replaced by the revised claim.",
  claim_a$statement,
  "Suitability for decisions about individual cases has not been established."
)
knitr::kable(data.frame(
  claim_id = c(claim_both$id, claim_a$id, claim_use$id),
  decision = c(decision_both$decision, decision_a$decision, decision_use$decision),
  action = c("revise", "retain", "withhold"), reported = reported_claims))
```

`met` holds relative to the listed required properties. It is not a certificate that every relevant property was
identified.

## Required properties, counterevidence, and context

Evidence on required properties decides the assessment (role `necessary_requirement`).
Counterevidence, such as an established coding error, counts against the claim once it is shown to affect the result
(role `potential_defeater` with materiality `materialized`).
Everything else is context: it is reported but never decides the claim (roles `graded_support` and
`descriptive_context`, both of which the paper calls context). In the package, leave `claim_consequence` at its
default for context records: a `graded_support` record with `claim_consequence = "unresolved"` leaves the claim
unresolved. A record list without any required property also leaves the claim unresolved.
A gate can be irrelevant to a claim that is otherwise in scope; a computation error makes evidence unavailable,
which leaves the claim unresolved rather than not met.

```{r status-boundaries}
irrelevant = csdg_evidence_record(
  "G4", FALSE, "descriptive_context",
  rationale = "Local-surrogate fidelity is not required by the global claim about A."
)
stopifnot(csdg_adjudicate_claim(c(evidence_a, list(irrelevant)))$decision == "met",
  csdg_adjudicate_claim(list(irrelevant))$decision == "unresolved")
# A deliberately simulated computation failure is an unavailable result, not counterevidence.
computation = tryCatch(stop("Illustrative computation error"), error = conditionMessage)
failed_evidence = csdg_evidence_record(
  "G2", TRUE, "necessary_requirement", availability = "unavailable",
  result_direction = "not_evaluated", rationale = computation
)
stopifnot(csdg_adjudicate_claim(list(failed_evidence))$decision == "unresolved")
```

A whole claim is `not_applicable` only if it is explicitly placed outside the evaluation with a rationale.

## Connect the record to a fitted mlr3 workflow

This separate example generates 180 observations, fits a small classification tree in each of three folds,
and computes held-out PFI. Each fold's model is fitted without the observations used to evaluate it, and the reported
values average over the fold-specific fits of the learner. The claim describes the learner's reliance on held-out data,
so it is declared at the functional level and, following Table 3 of the paper, does not require G1. The audit is
nevertheless asked to run the cross-validation listed under G1 (`run_gates`), because the stability diagnostic (G5)
uses its stored fold models.

```{r mlr3-audit}
set.seed(42)
dat = data.frame(x1 = rnorm(180), x2 = rnorm(180))
dat$y = factor(rbinom(180, 1, plogis(dat$x1 - 0.5 * dat$x2)), levels = c(0, 1))
task = as_task_classif(dat, target = "y", positive = "1")
learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 3L)
mlr_claim = csdg_claim(
  id = "tree_learner",
  statement = "The tree learner relies on both predictors.",
  claim_type = "global_explanation", target = "held-out log-loss increase under marginal permutation",
  unit = "generated observation", population = "the generated sample",
  analytic_distribution = "180 synthetic rows", model_scope = "cross_fitted_pipeline",
  semantics = "hypothetical_model_query",
  scientific_use = "describe the learner", claim_level = "functional",
  explanation_design = "three folds; 20 marginal permutations per predictor and fold"
)
mlr_measurement = csdg_measurement(
  outcome = "y", predictors = c("x1", "x2"), data_source = "generated here",
  sample_definition = "all 180 rows", unit = "generated observation",
  outcome_scale = "binary factor", missingness = "none", preprocessing = "none"
)
mlr_explanation = csdg_explanation(
  method_ids = "pfi", scope = "global", target = "held-out log loss",
  perturbation = list(type = "marginal", distribution = "assessment rows in each fold")
)
config = csdg_config(seed = 42L, resampling = list(folds = 3L, repeats = 1L),
  stability = list(pfi_repetitions = 20L, top_k = 2L))
audit = csdg_audit(task, learner, mlr_claim, mlr_measurement, mlr_explanation, config = config,
  run_gates = c("G0a", "G0b", "G1", "G2", "G5"))
report_card = csdg_report_card(audit)
stopifnot(!any(report_card$status == "error"))
knitr::kable(report_card[, c("gate_id", "status", "summary"), with = FALSE])
pfi = audit$artifacts$pfi$summary
knitr::kable(pfi[, c("feature_group", "mean_importance", "positive_fraction"), with = FALSE], digits = 4)
```

The report card uses the software's own labels for the diagnostics that the audit ran. G0b shows as unresolved
because no preprocessing check was passed to the audit, G1 and G5 because no criterion was passed, and gates outside the
software's plan show as `not_applicable`. These labels are not the supported, contradicted, or open status of a
required property. The high-level claim report deliberately leaves the assessment to explicit records, so a
successful audit alone never makes a claim met.

```{r claim-report}
claim_report = csdg_claim_report(audit)
knitr::kable(claim_report[, c("claim_id", "decision", "decision_basis"), with = FALSE])
stopifnot(all(claim_report$decision == "unresolved"))
```

The explicit assessment uses the audit's output and supplies what the audit lacked. The claim requires G0a, G0b, G2,
and G5; the researcher documents that the generated data involve no preprocessing (G0b), and the criterion for G5,
positive PFI in every fold beyond twice its Monte Carlo standard error, translates "relies on" into numbers and follows the paper's rule. The two predictors are independent, so permuted inputs are
realistic.

```{r mlr3-assessment}
fold_pfi = audit$artifacts$pfi$per_iteration
stopifnot(nrow(pfi) == 2L, all(fold_pfi$importance > 2 * fold_pfi$monte_carlo_se))
mlr_evidence = list(
  supported("G0a", "The claim and all six scope elements are stated.", "csdg_claim() records them.",
    "They fix the quantity, the learner, and the procedure."),
  supported("G0b", "The data match what the claim says.", "Generated data without missing values or preprocessing.",
    "The claim names only the generated sample."),
  supported("G2", "Held-out marginal permutation computes the log-loss increase named in the claim.",
    "Permutation within each held-out fold; the predictors are independent, so permuted inputs are realistic.",
    "The computation matches the quantity and procedure of the claim."),
  csdg_evidence_record(
    "G5", TRUE, "necessary_requirement", result_direction = "supports",
    claim_consequence = "retain_exact_claim",
    criterion = list(value = "PFI > 2 Monte Carlo standard errors in every fold", direction = "minimum"),
    criterion_source = "translated from the claim with the paper's stability rule",
    criterion_rationale = "\"Relies on\" means a positive held-out PFI beyond Monte Carlo error in each fold.",
    rationale = "Both predictors meet the criterion in every fold.",
    required_property = "Both predictors increase held-out log loss in every fold.",
    observation = sprintf("Positive PFI beyond twice its Monte Carlo standard error in all %d folds for both predictors.",
      pfi$n_iterations[1L]),
    relevance_to_proposition = "Folds vary the training and evaluation data while the learner and perturbation stay fixed.")
)
decision_mlr = csdg_adjudicate_claim(mlr_evidence, claim_applicable = TRUE)
decision_mlr$decision
stopifnot(identical(decision_mlr$decision, "met"))
```

## Export, reload, and verify the actual output

The explicit assessment file records the installed package version and, for all four claims, the evidence, decisions,
and reported claims, together with the revision relation. For the audit claim, it supersedes the audit bundle's own
claim report (unresolved). The file and a receipt are written next to the audit bundle, not into it, so the bundle's
checksum manifest stays valid; the receipt names the file, its SHA-256 checksum, the audit bundle, and all claim
identifiers. After reading the files back, the code checks the claim identifiers, the four decisions, the file
checksum, and the audit's unresolved claim report.

```{r export}
export_root = tempfile("csdg_walkthrough_")
dir.create(export_root)
audit_bundle = csdg_export(audit, path = export_root)
explicit = list(
  package_version = as.character(packageVersion("mlr3autoiml")),
  claims = list(claim_both, claim_a, claim_use, mlr_claim),
  evidence = list(evidence_both, evidence_a, evidence_use, mlr_evidence),
  decisions = list(decision_both, decision_a, decision_use, decision_mlr),
  reported_claims = c(reported_claims, mlr_claim$statement), relation = relation
)
assessment_path = file.path(export_root, "explicit_assessments.json")
jsonlite::write_json(explicit, assessment_path, auto_unbox = TRUE, pretty = TRUE, null = "null", na = "null")
receipt = list(
  assessment_file = basename(assessment_path),
  assessment_sha256 = digest::digest(file = assessment_path, algo = "sha256"),
  audit_claim_id = mlr_claim$id, audit_bundle = basename(audit_bundle),
  claim_ids = c(claim_both$id, claim_a$id, claim_use$id, mlr_claim$id)
)
jsonlite::write_json(receipt, file.path(export_root, "receipt.json"), auto_unbox = TRUE, pretty = TRUE)
read_back = jsonlite::read_json(assessment_path)
read_receipt = jsonlite::read_json(file.path(export_root, "receipt.json"))
stopifnot(identical(vapply(read_back$claims, `[[`, character(1L), "id"), receipt$claim_ids),
  identical(vapply(read_back$decisions, `[[`, character(1L), "decision"), c("not_met", "met", "unresolved", "met")),
  identical(read_receipt$assessment_sha256, digest::digest(file = assessment_path, algo = "sha256")),
  all(read.csv(file.path(audit_bundle, "claim_report.csv"))$decision == "unresolved"))
cat("Export and read-back verified for all four claims.\n")
# A caller can retain this synthetic-only bundle without changing the tutorial.
keep = Sys.getenv("CSDG_VIGNETTE_EXPORT_DIR")
if (nzchar(keep)) {
  dir.create(keep, recursive = TRUE, showWarnings = FALSE)
  stopifnot(all(file.copy(list.files(export_root, full.names = TRUE), keep, recursive = TRUE)))
}
```

## The applications in the paper

In SHILD, the claim that explanations of individual predictions can guide follow-up decisions remains unresolved,
because no criterion derived from the use and no evidence on the decisions exist. Three claims that answer other
questions are met: the local explanations are less accurate on new points than on the points used to fit them, and the
two accuracy measures disagree about where the explanations work best; for this rare outcome, local explanations
reproduce the model more accurately when they are fitted on the log-odds scale; and the model relies mainly on
self-rated health and depressed mood.

In PISA, the five most important predictors of the XGBoost model are the same for every plausible value (met), but
three similarly accurate models do not have the same five most important predictors (not met). The revised claim, that
all three models rely heavily on assessing credibility, but ridge regression relies far more on socioeconomic indices
than XGBoost and random forest do, is met. The paper's online supplement records the sources of variation and the
sensitivity analyses.

For further diagnostics, use the installed help pages (`?csdg_ale_bootstrap`, `?csdg_local_fidelity_audit`,
and `?csdg_importance_agreement`). They are optional analyses selected by a claim, not additional mandatory tutorial steps.

```{r session}
sessionInfo()
```
