---
title: "From a claim to evidence: a complete CSDG walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{From a claim to evidence: a complete CSDG walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = TRUE)
```

This guide follows the five steps of the accompanying article:
**write the claim, specify the scope, derive the required properties, evaluate the evidence, and decide and report**.
The input is a claim, a conclusion that a researcher intends to report. The output is its assessment (met, not met,
or unresolved) and a decision: a met claim is retained; a claim that is not met or unresolved is revised or withheld.
The package records the claim and its scope, derives the gate plan, computes diagnostics, and applies the decision
rule. The researcher sets the criteria, states the relevance of each observation, and enters established
counterevidence and unresolved threats.

All code runs on generated data; no SHILD or PISA files are needed. The first part works through the numerical
example of the article. The second part applies the steps to a fitted `mlr3` learner and shows an unresolved claim
with an unresolved threat.

```{r package, warning = FALSE, message = FALSE}
library(mlr3autoiml)
library(data.table)
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)
packageVersion("mlr3autoiml")
```

<style>
pre, pre code, pre > code.sourceCode { white-space: pre-wrap; overflow-wrap: anywhere; }
pre > code.sourceCode { display: block; }
pre > code.sourceCode > span { display: inline-block; max-width: 100%; white-space: pre-wrap; overflow-wrap: anywhere; }
table { width: 100%; table-layout: auto; }
th, td { overflow-wrap: anywhere; vertical-align: top; }
</style>

## Step 1: Write the claim

Suppose that two items, for example two wordings of the self-rated health question, carry the same information. For
exact arithmetic, let `Z` be standard normal, let both items and the outcome equal `Z`, and use squared error.
Model A is `f_A(x) = x1` and model B is `f_B(x) = x2`, both defined for all inputs and without error. The claim is:

> **Under marginal permutation, both models rely more on item 1 than on item 2.**

Reliance is measured by held-out permutation feature importance (PFI): the increase in expected squared error when
one item is permuted.

The record states that the claim was written before the values were computed (specified in advance).

## Step 2: Specify the scope

The scope states what the claim refers to, in six elements: the quantity (the estimand that the claim interprets),
the model, the procedure (the estimator that computes the quantity), the data, the meaning (model description,
population claim, or causal claim), and the use.

```{r claim}
claim_both = csdg_claim(
  id = "both_models",
  statement = "Under marginal permutation, both models rely more on item 1 than on item 2.",
  claim_type = "global_explanation",
  quantity = "marginal PFI with squared error: increase in expected squared error when one item is permuted",
  model = "several_models",
  procedure = paste(
    "each item permuted independently of the other item and the outcome;",
    "20 permutations per held-out fold of a 5-fold cross-validation"
  ),
  data = "population Z ~ N(0, 1), item 1 = item 2 = outcome = Z; 10,000 simulated observations",
  meaning = "model_description",
  use = "scientific description",
  unit = "simulated observation",
  population = "Z standard normal",
  provenance = list(origin = "specified_before_results", date = "2026-09-26",
    time_basis = "date of the example", selection_basis = "written before the PFI values were computed",
    evidence_ids = character())
)
knitr::kable(data.frame(element = names(claim_both$scope), content = unlist(claim_both$scope)),
  row.names = FALSE)
```

Marginal permutation creates input combinations that never occur in the data: a permuted input has unequal items.
A model description may cover such inputs if its scope says so, as must its reported wording. The qualifier "under
marginal permutation" is therefore part of the claim; without it, the claim would be broader than its scope.

```{r scope}
measurement = csdg_measurement(
  outcome = "y", predictors = c("item_1", "item_2"),
  data_source = "simulated from the specified population",
  sample_definition = "all simulated observations; no exclusions",
  unit = "simulated observation", outcome_scale = "continuous, variance one",
  missingness = "none", preprocessing = "none"
)
groups = list(item_1 = "item_1", item_2 = "item_2", both_items = c("item_1", "item_2"))
explanation = csdg_explanation(
  method_ids = "pfi", scope = "global", target = "prediction", feature_groups = groups,
  perturbation = list(type = "marginal", distribution = "held-out fold")
)
```

## Step 3: Derive the required properties

A gate is a generic evidence question, such as "Does the result hold for all models the claim covers?" Applied to a
claim, it becomes a required property, such as "item 1 has the larger PFI in model B." The package derives the gate
plan from the column "Required if the claim ..." of the gate registry (Table 3 of the article).

```{r plan}
plan = csdg_gate_plan(claim_both, measurement, explanation)
knitr::kable(plan[, .(gate_id, gate_name, required, plan_role, trigger)])
stopifnot(identical(plan[required == TRUE, gate_id], c("G0a", "G0b", "G2", "G5", "G6a")))
```

Besides G0a and G0b, the claim requires G2 (it interprets PFI values), G5 (it states an ordering), and G6a (it covers
two models). Held-out performance (G1) is context: a model description holds whatever the model's accuracy.

## Step 4: Evaluate the evidence

In model A, shuffling item 1 replaces the prediction `Z` with an independent draw `Z'` from the same distribution, so
the expected squared error rises from 0 to `E[(Z - Z')^2] = Var(Z) + Var(Z') = 2`; shuffling item 2 leaves the
prediction unchanged, so its PFI is 0. In model B, the values are reversed. These are exact population values. The
package computes them approximately from 10,000 simulated observations: each model is a linear model of one item,
which reproduces the outcome exactly in every fold.

```{r compute}
set.seed(260926)
n = 10000L
z = rnorm(n)
task = as_task_regr(data.frame(item_1 = z, item_2 = z, y = z), target = "y", id = "numerical_example")
model_using = function(item) {
  as_learner(po("select", selector = selector_name(item), id = paste0("use_", item)) %>>% lrn("regr.lm"))
}
folds = rsmp("cv", folds = 5L)
fits = list(
  A = csdg_resample(task, model_using("item_1"), folds, measures = msrs("regr.mse"), seed = 1L),
  B = csdg_resample(task, model_using("item_2"), folds, measures = msrs("regr.mse"), seed = 1L)
)
pfi = lapply(fits, csdg_fold_pfi, feature_groups = groups, loss = "mse", repetitions = 20L, seed = 2L)
marginal = dcast(rbindlist(lapply(names(pfi), function(model) {
  pfi[[model]]$summary[, .(model = model, feature_group, mean_importance)]
})), model ~ feature_group, value.var = "mean_importance")
knitr::kable(marginal, digits = 3,
  caption = "Held-out PFI with squared error; the exact values are (2, 0) in A and (0, 2) in B.")
stopifnot(marginal[model == "A", item_1 > item_2], marginal[model == "B", item_1 < item_2])
```

The Monte Carlo rule decides whether a difference is beyond the error due to random permutation: its absolute value
must exceed the .975 quantile of the t distribution with R - 1 degrees of freedom times its Monte Carlo standard
error, where R is the number of permutations. This rule addresses permutation noise; the interval below describes
sampling uncertainty. In model A, the difference between the items exceeds it in every fold. The corrected resampled t interval of the
learner-level mean difference is reported descriptively.

```{r monte-carlo}
mc_a = csdg_pfi_mc_difference(pfi$A, "item_1", "item_2")
knitr::kable(mc_a[, .(iteration, estimate, se, threshold, beyond_monte_carlo_error)], digits = 4)
stopifnot(all(mc_a$beyond_monte_carlo_error))
interval_a = csdg_learner_pfi_interval(pfi$A, n_train = 8000, feature_group = "item_1", minus = "item_2")
knitr::kable(interval_a[, .(estimate, se, lower, upper, df)], digits = 4)
```

Other versions of PFI answer other questions. Conditional PFI draws the permuted item from its distribution given the
other item and so keeps the inputs realistic; because the items are equal, the draw equals the original value, and
conditional PFI is 0 for both items in both models. Here, permuting within strata of `Z` implements this.
Grouped PFI, which shuffles both items together, is about 2 in both models (the column `both_items` above).

```{r conditional}
conditional = lapply(fits, csdg_fold_pfi, feature_groups = groups[1:2], loss = "mse", repetitions = 5L,
  strata = z, seed = 3L)
conditional_table = dcast(rbindlist(lapply(names(conditional), function(model) {
  conditional[[model]]$summary[, .(model = model, feature_group, mean_importance)]
})), model ~ feature_group, value.var = "mean_importance")
knitr::kable(conditional_table, digits = 3, caption = "Conditional PFI.")
stopifnot(all(abs(as.matrix(conditional_table[, .(item_1, item_2)])) < 1e-12))
```

For each required property, the record contains four entries: the property, the criterion that decides whether it
holds, the observation, and its relevance. Judged against its criterion, each required property is supported,
contradicted, or open; `supported()`, `contradicted()`, and `open_status()` return these statuses.
`csdg_claim_record_template()` returns the blank claim record of the article with all its fields.

```{r evidence-supported}
property = function(gate_id, status, required_property, observation, relevance, criterion = NULL) {
  csdg_evidence_record(
    gate_id, TRUE, "required_property", status = status,
    criterion = if (!is.null(criterion)) list(value = criterion, direction = "qualitative"),
    criterion_source = if (!is.null(criterion)) "translated from the claim",
    criterion_rationale = if (!is.null(criterion)) required_property,
    rationale = observation, required_property = required_property, observation = observation,
    relevance_to_proposition = relevance
  )
}
foundation = list(
  property("G0a", "supported", "The claim and its six scope elements are stated.",
    "csdg_claim() records all six elements.", "They fix the quantity, the models, and the procedure."),
  property("G0b", "supported", "The data cover what the claim names.",
    "The population is defined exactly; there is no measurement error or preprocessing.",
    "The claim names no construct and no other population."),
  property("G2", "supported", "Marginal permutation computes the PFI that the claim names.",
    "Squared error and marginal permutation; both models are defined for all inputs.",
    "The claim names the perturbation, so the unrealistic permuted inputs are part of what it describes."),
  property("G5", "supported", "The ordering persists when the quantity is estimated again.",
    "Exact values do not vary.", "The same quantity is computed again.")
)
```

The decisive property is G6a: only the model differs between the two computations, the claim covers both models, and
model B reverses the ordering. The status records the researcher's judgment: the researcher applies the criterion to
the numbers.

```{r evidence-both}
g6a = property("G6a", "contradicted", "Item 1 has the larger marginal PFI in A and in B.",
  "PFI of item 1 versus item 2: 2 versus 0 in A, 0 versus 2 in B.",
  "Only the model differs between the two computations, and the claim covers both models.",
  criterion = "strict ordering in both models")
evidence_both = c(foundation, list(g6a))
```

## Step 5: Decide and report

The assessment follows a fixed rule: a claim is not met if at least one required property is contradicted; otherwise
it is unresolved if at least one is open; otherwise it is met. Favorable results never offset a contradicted
property. With the plan, a required gate without evidence would be open.

```{r decide-both}
assessment_both = csdg_adjudicate_claim(evidence_both, claim_applicable = TRUE, plan = plan)
assessment_both
stopifnot(identical(assessment_both$assessment, "not_met"),
  identical(assessment_both$decision_options, c("revise", "withhold")))
```

The original claim is not met; it faces contrary evidence and is revised. The revised claim, restricted to model A,
is a new claim with its own record and assessment. It was written after the results were seen, so its assessment is
exploratory.

```{r revised-claim}
claim_a = csdg_claim_revision(
  claim_both, id = "model_a", claim_version = "C1", revision_relation = "narrower",
  statement = "Under marginal permutation, model A relies more on item 1 than on item 2.",
  model = "fitted_model",
  provenance = list(origin = "retrospective_exploratory", date = "2026-09-26",
    time_basis = "date of the example",
    selection_basis = "restriction chosen after the comparison of A and B", evidence_ids = "both_models_G6a")
)
plan_a = csdg_gate_plan(claim_a, measurement, explanation)
stopifnot(identical(plan_a[required == TRUE, gate_id], c("G0a", "G0b", "G2", "G5")))
g2_a = property("G2", "supported",
  "Marginal PFI with squared error in model A is larger for item 1, and the wording names the perturbation.",
  "2 versus 0; conditional PFI 0 versus 0.",
  "The computation uses the model, perturbation, and loss that the claim names.",
  criterion = "strict ordering of the exact values; qualifier in the wording")
evidence_a = c(foundation[c(1L, 2L, 4L)], list(g2_a))
assessment_a = csdg_adjudicate_claim(evidence_a, claim_applicable = TRUE, plan = plan_a)
assessment_a$assessment
stopifnot(identical(assessment_a$assessment, "met"), identical(assessment_a$decision_options, "retain"))
```

The revised claim is met (exploratory) and retained. Narrowing a claim can reuse the evidence for the narrower scope.
The record compares the revised claim with its predecessor in two ways: for each scope element, whether the new scope
is the same, narrower, broader, or incomparable; and, separately, whether the revised claim follows from its
predecessor. Here it does: a claim about both models is universal, so it implies the claim about model A (a logical
weakening). The package records this argument as the researcher states it.

```{r relation}
relation = csdg_claim_relation(
  claim_both, claim_a,
  coordinate_relations = c(quantity = "same", model = "narrower", procedure = "same", data = "same",
    meaning = "same", use = "same"),
  rationale = "The claim about models A and B is restricted to model A; everything else is unchanged.",
  proposition_relation = "logical_weakening",
  proposition_rationale = "A claim that holds for every model in {A, B} holds for A.",
  revision_kind = "logical_weakening"
)
knitr::kable(relation$scope)
```

The assessment of a met claim refers to the required properties listed in its record.

## An unresolved claim: an unresolved threat

The second part applies the steps to a fitted `mlr3` learner. The generated data contain 180 observations and ten
candidate predictors; the two predictors of the task were chosen by screening all ten on the full sample, before the
cross-validation. The claim uses the predictions, so held-out performance (G1) is required.

```{r threat-data}
set.seed(42)
candidates = as.data.frame(matrix(rnorm(180 * 10), ncol = 10, dimnames = list(NULL, paste0("x", 1:10))))
outcome = factor(rbinom(180, 1, plogis(1.5 * candidates$x1 - candidates$x2)), levels = c(0, 1))
screening = sapply(candidates, function(x) abs(cor(x, as.numeric(outcome))))
selected = names(sort(screening, decreasing = TRUE))[1:2]
selected
dat = data.frame(candidates[selected], y = outcome)
task_screened = as_task_classif(dat, target = "y", positive = "1", id = "screened")
learner = lrn("classif.log_reg", predict_type = "prob")
claim_predict = csdg_claim(
  id = "learner_predicts",
  statement = "The logistic regression learner predicts the outcome better than a model without predictors.",
  claim_type = "predictive_performance",
  quantity = "held-out AUC of the learner and of a model without predictors",
  model = "learner",
  procedure = "3-fold cross-validation; the learner is refitted in each training fold",
  data = "180 generated observations; two predictors screened from ten candidates on the full sample",
  meaning = "model_description", use = "scientific description",
  unit = "generated observation", population = "the generated sample"
)
measurement_screened = csdg_measurement(
  outcome = "y", predictors = selected, data_source = "generated in this vignette",
  sample_definition = "all 180 rows", unit = "generated observation", outcome_scale = "binary factor",
  missingness = "none", preprocessing = "screening of ten candidates by correlation with the outcome",
  verification = list(status = "author_reviewed", artifact = "the generating code of this vignette")
)
csdg_gate_plan(claim_predict, measurement_screened)[required == TRUE, gate_id]
```

The criterion for G1 translates "better than a model without predictors" into a margin with a source and a rationale.

```{r threat-audit}
config = csdg_config(
  seed = 42L, resampling = list(folds = 3L, repeats = 1L),
  performance = list(minimum_baseline_improvement = 0.05),
  criteria = list(performance.minimum_baseline_improvement = list(
    source = "translated from the claim",
    rationale = "\"Better\" means a held-out AUC at least 0.05 above that of the model without predictors."
  ))
)
audit_predict = csdg_audit(task_screened, learner, claim_predict, measurement_screened, config = config)
knitr::kable(csdg_report_card(audit_predict)[required == TRUE, .(gate_id, gate_name, status, summary)])
```

The audit supports the G1 property under its criterion. But the predictors were selected on the same data outside
the cross-validation, so the held-out values could be optimistic: a specific and plausible, but unquantified,
problem. It is an unresolved threat to the G1 property, which is therefore open until a nested analysis quantifies
the optimism.

```{r threat}
threat = csdg_evidence_record(
  "G1", TRUE, "unresolved_threat",
  required_property = "Held-out performance is not inflated by predictor selection.",
  observation = "The predictors were selected on the same data outside the cross-validation.",
  rationale = "The optimism was not quantified; a nested analysis that repeats the screening in each training fold would resolve it."
)
report_predict = csdg_claim_report(audit_predict, evidence = list(threat))
knitr::kable(report_predict[, .(claim_id, assessment, decision_options, supported_gates, open_gates, threat_gates)])
stopifnot(identical(report_predict$assessment, "unresolved"))
```

The claim is unresolved. It lacks evidence, which a nested analysis or new data could supply; until then, it is
withheld and, where it matters, reported as not established.

## The same learner: a met claim with the audit

A model description of the same learner, "The logistic regression learner relies on both selected predictors,"
requires G2 and G5;
held-out performance is context, and the predictor selection bears on it only through performance, which this claim
does not interpret. The audit computes the diagnostics: dependence and observed support for G2, and held-out PFI in
every fold for G5. Whether the procedure computes the quantity that the claim names is recorded by the researcher;
until then, G2 is open.

```{r mlr3-audit}
claim_relies = csdg_claim(
  id = "learner_relies",
  statement = "The logistic regression learner relies on both selected predictors.",
  claim_type = "global_explanation",
  quantity = "held-out log-loss increase under marginal permutation of each predictor",
  model = "learner",
  procedure = "3-fold cross-validation; 20 marginal permutations per predictor and held-out fold",
  data = "180 generated observations with the two screened predictors",
  meaning = "model_description", use = "scientific description",
  unit = "generated observation", population = "the generated sample"
)
explanation_screened = csdg_explanation(method_ids = "pfi", scope = "global", target = "held-out log loss",
  perturbation = list(type = "marginal", distribution = "assessment rows in each fold"))
config_relies = csdg_config(seed = 42L, resampling = list(folds = 3L, repeats = 1L),
  stability = list(pfi_repetitions = 20L, top_k = 2L))
audit = csdg_audit(task_screened, learner, claim_relies, measurement_screened, explanation_screened, config = config_relies)
report_card = csdg_report_card(audit)
stopifnot(!any(report_card$status == "error"))
stopifnot(identical(report_card[gate_id == "G1", status], "context"))
knitr::kable(report_card[, .(gate_id, gate_name, required, evidence_role, status, diagnostic_status)])
pfi_learner = audit$artifacts$pfi$per_iteration
knitr::kable(pfi_learner[, .(iteration, feature_group, importance, monte_carlo_se)], digits = 4)
```

Held-out performance (G1) is computed and reported as context: it has no property status, and its computed result
(`diagnostic_status`) never changes the assessment.
The criterion for G5, positive PFI in every fold beyond twice its Monte Carlo standard error, translates "relies on"
into numbers. The two predictors were generated independently, so permuted inputs are realistic (G2).

```{r mlr3-assessment}
stopifnot(all(pfi_learner$importance > 2 * pfi_learner$monte_carlo_se))
g2_learner = property("G2", "supported",
  "Held-out marginal permutation computes the log-loss increase that the claim names.",
  "Permutation within each held-out fold; the predictors are independent, so permuted inputs are realistic.",
  "The computation uses the learner, perturbation, and loss that the claim names.")
g5_learner = property("G5", "supported", "Both predictors increase held-out log loss in every fold.",
  sprintf("PFI above twice its Monte Carlo standard error in all %d folds for both predictors.",
    length(unique(pfi_learner$iteration))),
  "Folds vary the training and evaluation data while the learner and perturbation stay fixed.",
  criterion = "PFI > 2 Monte Carlo standard errors in every fold")
report_relies = csdg_claim_report(audit, evidence = list(g2_learner, g5_learner))
knitr::kable(report_relies[, .(claim_id, assessment, decision_options, supported_gates, open_gates)])
stopifnot(identical(report_relies$assessment, "met"))
```

## Counterevidence, context, errors, and claims outside the evaluation

Established counterevidence is a demonstrated problem, such as a coding error shown to affect the result. It
contradicts the property it bears on, so an otherwise met claim is not met. Context is reported but never changes
the assessment. A computation error leaves the property open, not contradicted, and a whole claim is
`not_applicable` only if it is explicitly placed outside the evaluation, with a rationale.

```{r other-roles}
coding_error = csdg_evidence_record(
  "G2", TRUE, "established_counterevidence",
  required_property = "Held-out marginal permutation computes the log-loss increase that the claim names.",
  observation = "Illustration: the two predictor columns were swapped before the PFI computation.",
  relevance_to_proposition = "The swap changes which predictor each reported value refers to.",
  rationale = "A coding error shown to affect the result."
)
grouped = csdg_evidence_record("G5", TRUE, "context", result_direction = "descriptive",
  rationale = "Grouped PFI of both predictors is reported for comparison; it does not bear on each predictor.")
computation = tryCatch(stop("Illustrative computation error"), error = conditionMessage)
failed = csdg_evidence_record("G2", TRUE, "required_property", status = "open", availability = "unavailable",
  rationale = computation)
stopifnot(
  identical(csdg_claim_report(audit, evidence = list(g2_learner, g5_learner, coding_error))$assessment, "not_met"),
  identical(csdg_claim_report(audit, evidence = list(g2_learner, g5_learner, grouped))$assessment, "met"),
  identical(csdg_claim_report(audit, evidence = list(failed, g5_learner))$assessment, "unresolved"),
  identical(csdg_adjudicate_claim(list(), claim_applicable = FALSE,
    applicability_rationale = "The claim is outside this evaluation.")$assessment, "not_applicable")
)
```

## Export, reload, and verify

The audit bundle contains the claim report of the audit claim, assessed by the decision rule from the audit
diagnostics alone: without the researcher's records, G2 and G5 are open, so the claim is unresolved. The assessments
of all four claims, with the researcher's records, are written to a separate file next to the bundle and read back.

```{r export}
export_root = tempfile("csdg_walkthrough_")
dir.create(export_root)
audit_bundle = csdg_export(audit, path = export_root)
explicit = list(
  package_version = as.character(packageVersion("mlr3autoiml")),
  claims = list(claim_both, claim_a, claim_predict, claim_relies),
  evidence = list(evidence_both, evidence_a, list(threat), list(g2_learner, g5_learner)),
  assessments = list(assessment_both$assessment, assessment_a$assessment, report_predict$assessment,
    report_relies$assessment),
  decisions = c("revise", "retain", "withhold", "retain"),
  reported_claims = c("Revised to the claim about model A.", claim_a$statement,
    "Withheld: not established.", claim_relies$statement),
  relation = relation
)
assessment_path = file.path(export_root, "explicit_assessments.json")
jsonlite::write_json(explicit, assessment_path, auto_unbox = TRUE, pretty = TRUE, null = "null", na = "null")
read_back = jsonlite::read_json(assessment_path)
bundle_report = read.csv(file.path(audit_bundle, "claim_report.csv"))
stopifnot(
  identical(vapply(read_back$claims, `[[`, character(1L), "id"),
    c(claim_both$id, claim_a$id, claim_predict$id, claim_relies$id)),
  identical(unlist(read_back$assessments), c("not_met", "met", "unresolved", "met")),
  identical(unlist(read_back$decisions), c("revise", "retain", "withhold", "retain")),
  identical(bundle_report$assessment, "unresolved"), identical(bundle_report$open_gates, "G2, G5"))
cat("Export and read-back verified for all four claims.\n")
# A caller can retain this synthetic-only bundle without changing the tutorial.
keep = Sys.getenv("CSDG_VIGNETTE_EXPORT_DIR")
if (nzchar(keep)) {
  dir.create(keep, recursive = TRUE, showWarnings = FALSE)
  stopifnot(all(file.copy(list.files(export_root, full.names = TRUE), keep, recursive = TRUE)))
}
```

## The applications in the article

In SHILD, the claim that XGBoost relies mainly on self-rated health and depressed mood is met (exploratory) and
retained, as are the claims that apparent fidelity overstates how accurately the local surrogates reproduce the model
on new points and that, for this rare outcome, surrogates fitted on the log-odds scale reproduce it more accurately.
The claim that explanations of individual predictions can guide follow-up decisions is not met, because the stability
property (G5) is contradicted, and is withheld. The revised population claim about the information that the two items
carry is unresolved, because agreement among similarly accurate learners (G6a) is open, and is withheld.

In PISA, the claim that the five most important predictors of XGBoost are the same for every plausible value is met
(a sanity check). The claim that three similarly accurate learners have the same five most important predictors is
not met and is revised; the claim that ridge regression relies far more on socioeconomic indices than XGBoost and
random forest do is not met and is withheld. In the confirmatory stage, four of five claims are met. The article
and its supplement report the records, criteria, and numbers.

For further diagnostics, see the help pages `?csdg_audit`, `?csdg_local_fidelity_audit` (the fitted local surrogate
is described in `?csdg_diagnostics`, section "Local surrogate"), `?csdg_learner_pfi_interval`, and
`?csdg_importance_agreement`.

```{r session-info}
sessionInfo()
```
