---
title: "Claim-Scoped Diagnostic Gates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Claim-Scoped Diagnostic Gates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why claims come first

An explanation method does not determine what its output can support. A global
importance ranking for one fitted model, for example, does not by itself show
that the ranking survives across resamples, near-equivalent models, cohorts,
or plausible setting shifts. CSDG therefore starts from an explicit claim and
assigns diagnostic gates to that claim.

```{r cards}
library(mlr3)
library(mlr3autoiml)

claim = csdg_claim(
  id = "example_global_pattern",
  statement = paste(
    "The selected pipeline predicts the observed binary outcome adequately",
    "and supports a stable global importance description in the analytic sample."
  ),
  claim_type = c(
    "predictive_performance", "calibration", "global_explanation"
  ),
  target = "binary outcome at the observed assessment",
  unit = "participant",
  population = "the worked-example population",
  analytic_distribution = "the generated analytic sample",
  model_scope = "selected_model",
  setting_scope = "analytic_sample"
)

measurement = csdg_measurement(
  outcome = "y",
  predictors = c("x1", "x2", "group"),
  data_source = "generated worked example",
  sample_definition = "all generated rows",
  unit = "participant",
  outcome_scale = "binary factor, positive class 1",
  missingness = "no missing values",
  preprocessing = "rpart pipeline refitted within each training split",
  verification = list(
    status = "author_reviewed",
    artifact = "vignette source"
  )
)

explanation = csdg_explanation(
  method_ids = "pfi",
  scope = "global",
  target = "positive-class predicted probability",
  perturbation = list(
    type = "marginal",
    distribution = "held-out analytic-sample folds"
  )
)

csdg_gate_plan(claim, measurement, explanation)
```

## Fit and diagnose a complete pipeline

Every learned preprocessing operation should be embedded in the learner so
that it is refitted inside each training split. The small example below has no
missingness or learned preprocessing, but the same rule applies to a graph
learner containing imputation, encoding, selection, and tuning.

```{r audit}
set.seed(42)
n = 180
dat = data.frame(
  x1 = rnorm(n),
  x2 = rnorm(n),
  group = factor(sample(letters[1:3], n, replace = TRUE))
)
eta = 0.8 * dat$x1 - 0.5 * dat$x2 + 0.3 * (dat$group == "b")
dat$y = factor(rbinom(n, 1, plogis(eta)), levels = c(0, 1))

task = as_task_classif(dat, target = "y", positive = "1")
learner = lrn("classif.rpart", predict_type = "prob")

config = csdg_config(
  seed = 42L,
  resampling = list(folds = 3L, repeats = 1L),
  stability = list(
    pfi_repetitions = 2L,
    top_k = 2L,
    min_top_k_overlap = 0.25
  )
)

result = csdg_audit(
  task = task,
  learner = learner,
  claim = claim,
  measurement = measurement,
  explanation = explanation,
  config = config
)

csdg_report_card(result)
csdg_claim_report(result)
```

## Adding model-refit uncertainty to ALE curves

A one-dimensional ALE curve from a final full-sample fit describes one fitted model.
When the claim also requires evidence that the curve persists across resampled analytic data, use
`csdg_ale_bootstrap()` with the independent sampling unit and the complete learner pipeline.
The returned percentile intervals are pointwise, not simultaneous, and remain conditional on the fixed learner
specification and analysis design.

```{r ale-bootstrap, eval=FALSE}
fitted_learner = learner$clone(deep = TRUE)
fitted_learner$train(task)

ale_intervals = csdg_ale_bootstrap(
  task = task,
  learner = fitted_learner,
  features = c("x1", "x2"),
  sample_n = 180L,
  ale_bins = 10L,
  replicates = 200L,
  resampling_unit = "row",
  min_interval_n = 10L,
  seed = 42L
)

csdg_plot_data(ale_intervals$ale_1d, type = "ale", style = "monochrome")
```

For clustered observations, set `resampling_unit = "cluster"`, supply one cluster label per task row, and use
`strata` when clusters must be sampled independently within prespecified strata.

## Reading status labels

`not_applicable` means that the prespecified claim does not invoke a gate.
`unresolved` means that a required component, criterion, or scope check is missing or cannot yet be adjudicated.
`not_met` records a prespecified diagnostic failure. `error` preserves a computational failure rather than silently
converting it into support. `met` means only that the evidence required by the declared gate was available and met
the declared criterion.

None of these labels converts a predictive association into a causal effect.
Nor do they establish population transport, fairness, user comprehension, or
real-world utility unless those components were explicitly designed and
supplied.

## Exporting the audit trail

```{r export, eval=FALSE}
bundle = csdg_export(result, path = "results")
bundle
```

The bundle contains machine-readable cards, the gate plan, report and claim
cards, diagnostic tables, a serialized result, session information, an
uncertainty-scope note, and file checksums.
