---
title: "Checking what an explanation shows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking what an explanation shows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
run = all(vapply(c("mlr3learners", "xgboost", "glmnet"), requireNamespace, logical(1L), quietly = TRUE))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5, eval = run)
num = function(x, digits = 3L) if (run) formatC(signif(x, digits), digits = digits, format = "fg", flag = "#") else "?"
txt = function(x) if (run) x else "?"
```

An explanation of a fitted model, such as permutation feature importance (PFI), often suggests a conclusion: "the
model relies mainly on these two predictors." This vignette checks such conclusions, called claims, in the order of
an ordinary analysis: fit learners, compute the explanation and look at it, state the claim that the result suggests,
and let the package compute the checks of Claim-Scoped Diagnostic Gates (CSDG) and the assessment. The package
writes the claim, its scope, the criteria, and the observations from the computations.
`r if (!run) "The code is not run because the suggested packages mlr3learners, xgboost, or glmnet are not installed."`

```{r packages, message = FALSE, warning = FALSE}
library(mlr3autoiml)
library(mlr3)
library(mlr3learners)
library(data.table)
```

## Data

The simulated survey has 1,000 respondents, 12 items, and a binary outcome, `low_wellbeing`. Two scale scores
(`low_mood`, `fatigue`) predict it strongly; `worry_a` and `worry_b` are two wordings of the same 5-point worry item,
with nearly identical answers; `poor_sleep`, `stress`, and `loneliness` have weak effects; the other five items have
none.

```{r data}
simulate_survey = function(n, seed) {
  set.seed(seed)
  items = c("low_mood", "fatigue", "worry_a", "worry_b", "poor_sleep", "stress", "loneliness", "inactivity",
    "appetite", "concentration", "support", "optimism")
  x = as.data.table(matrix(rnorm(n * 12), n, 12, dimnames = list(NULL, items)))
  worry = rnorm(n)
  # two wordings of the same 5-point worry item; 2% of respondents answer them one category apart
  x[, worry_a := as.integer(cut(worry + rnorm(n, sd = 0.3), c(-Inf, -1.2, -0.4, 0.4, 1.2, Inf)))]
  x[, worry_b := pmin(5L, pmax(1L, worry_a + ifelse(runif(n) < 0.02, sample(c(-1L, 1L), n, TRUE), 0L)))]
  eta = -0.3 + 1.2 * x$low_mood + 1.1 * x$fatigue + 0.7 * worry + 0.3 * x$poor_sleep + 0.2 * x$stress +
    0.15 * x$loneliness
  x[, low_wellbeing := factor(rbinom(n, 1, plogis(eta)), levels = c(0, 1))]
  as_task_classif(x, target = "low_wellbeing", positive = "1", id = "survey")
}
task = simulate_survey(1000L, seed = 3L)
```

## Fit and look

`csdg_fit()` fits XGBoost and ridge logistic regression with `mlr3::benchmark()` on the same ten cross-validation
splits, together with a model without predictors. The loss, log loss by default, is used for the fold scores, for
PFI, and for the improvement over the model without predictors.

```{r fit}
learners = list(
  xgboost = lrn("classif.xgboost", nrounds = 100, eta = 0.1, max_depth = 3, predict_type = "prob"),
  ridge = lrn("classif.cv_glmnet", alpha = 0, nfolds = 5, predict_type = "prob")
)
fits = csdg_fit(task, learners, folds = 10, seed = 1)
fits
```

`csdg_importance()` computes held-out PFI, the increase in held-out log loss when one predictor is permuted, in each
fold with 20 permutations.

```{r importance}
imp = csdg_importance(fits, repetitions = 20, seed = 2)
imp
plot(imp)
```

The printout reports the improvement of each learner over the model without predictors, here
`r num(imp$improvement[learner == "xgboost", improvement])` log-loss units for XGBoost. PFI values are not parts of
this improvement, so PFI relative to the improvement can exceed 100%. In the plot, the bars are averages over folds,
the points are folds, and the dashed line marks 1% of the improvement, the default minimum importance.
`summary(imp)` returns the table with corrected 95% intervals for a results table.

## Claim 1: XGBoost relies mainly on two items

The result suggests that XGBoost relies mainly on `low_mood` and `fatigue`. `claim_relies_mainly()` states this as a
claim: each of the two predictors with the largest average PFI has at least twice the PFI of every other predictor.
`csdg_check()` computes the properties that the claim requires.

```{r claim-1}
claim1 = claim_relies_mainly(imp, k = 2, factor = 2, learners = "xgboost")
claim1
chk1 = csdg_check(claim1, imp)
chk1
```

Each row is one property with its gate, check, and status. Under G2, the content row compares the averages over
folds; the margin is the PFI of each named item minus twice the PFI of the largest other item, and it must hold
beyond Monte Carlo error and with the lower bound of its corrected resampled t interval. The minimum importance
compares each named item with 1% of the improvement, and the procedure row records that the wording names marginal
PFI. Under G5, the selection is repeated in each fold: `r txt(chk1$details$stability$xgboost$n_holds)` of
`r txt(fits$K)` folds reproduce the claim. The result of ridge logistic regression (G6a) and held-out performance
(G1) are context. `print(chk1, details = TRUE)` adds each criterion and its relevance; `?csdg_check` documents the
criteria and their sources.

```{r assess-1}
res1 = csdg_assess(chk1)
res1
```

```{r assert-1, include = FALSE}
stopifnot(identical(res1$assessment, "met"))
```

The claim is `r txt(res1$label)`. It is exploratory because it was formulated after the explanation was inspected;
the data that suggested it cannot also test it.

## Claim 2: one wording matters more than the other

XGBoost gives the two wordings of the worry item different PFI. Given the pair, `claim_order()` orders it by the
average PFI and states that XGBoost relies more on the first wording than on the second.

```{r claim-2}
claim2 = claim_order(imp, c("worry_a", "worry_b"), learners = "xgboost")
chk2_marginal = csdg_check(claim2, imp)
csdg_assess(chk2_marginal)
```

The two wordings correlate at
r = `r txt(sub("^0", "", num(cor(task$data()$worry_a, task$data()$worry_b), 2L)))`. Marginal PFI can shift
importance between correlated predictors, so without a comparison that accounts for the correlation, the claim faces
an unresolved threat under G2 and is `r txt(csdg_assess(chk2_marginal)$label)`. Conditional PFI permutes one wording
within the answers to the other, so the permuted inputs stay realistic. Passing the existing result to
`csdg_importance()` adds it; the marginal values do not change.

```{r claim-2-conditional}
imp2 = csdg_importance(imp, conditional = list(worry_a = "worry_b", worry_b = "worry_a"))
plot(imp2, features = c("worry_a", "worry_b"))
res2 = csdg_assess(csdg_check(claim2, imp2))
res2
```

```{r assert-2, include = FALSE}
stopifnot(res2$assessment %in% c("not_met", "unresolved"))
```

Once one wording is known, the other adds
`r num(imp2$summary[learner == "xgboost" & type == "conditional" & feature_group == claim2$first, mean_importance], 2L)`
(`r txt(claim2$first)`) and
`r num(imp2$summary[learner == "xgboost" & type == "conditional" & feature_group == claim2$second, mean_importance], 2L)`
(`r txt(claim2$second)`) log-loss units. The claim is `r txt(res2$label)` (decisive:
`r txt(as.data.table(res2)$decisive_checks)`): conditional PFI does not establish that either wording adds
information beyond the other. The check also reports the same comparison for ridge logistic regression as context
(G6a). With `marginal_only = TRUE`, the claim would describe only
marginal permutation; which wording XGBoost uses can change with the sample.

## A direction

`csdg_effect()` computes the accumulated local effect (ALE) of predictors in the held-out folds and its change
between the 10th and the 90th percentile. After looking at it, `claim_direction()` states the sign.

```{r direction}
eff = csdg_effect(fits, c("poor_sleep", "stress"))
eff
claim3 = claim_direction(eff, "poor_sleep", learners = "xgboost")
res3 = csdg_assess(csdg_check(claim3, eff))
res3
```

This claim is `r txt(res3$label)`.

## Claim 1 on new data

`csdg_confirm()` applies claim 1 with the criteria of `chk1` to a new sample from the same population. Given a new
task, it fits the same learners with the same resampling and seeds and computes PFI with the same settings; the two
predictors are not selected again, and the criteria were fixed before the new data were analyzed. Rows that also
occur in the data of the check are an error.

```{r confirm}
task_new = simulate_survey(1000L, seed = 1003L)
conf1 = csdg_confirm(chk1, task_new)
conf1
```

```{r assert-3, include = FALSE}
stopifnot(identical(conf1$assessment, "met"))
```

Claim 1 is `r txt(conf1$label)`.

## Constructs and judgments

A claim that names constructs rather than the analyzed variables needs validity evidence for them (G0b), which the
package cannot compute. `refers_to = "constructs"` marks such a claim, and `csdg_judge()` enters the evidence.

```{r judge}
claim1c = claim_relies_mainly(imp, k = 2, learners = "xgboost", refers_to = "constructs")
judged = csdg_judge("G0b", supported(), note = "Both scales: omega above .85 and scalar invariance (Table S2).")
csdg_assess(csdg_check(claim1c, imp, judgments = judged))$label
```

## The records

The printouts are summaries. `as.data.table(chk1)` holds the full property, criterion, observation, and relevance
texts; `chk1$records` are the evidence records built with `csdg_evidence_record()`, and `res1$adjudication` is the
result of `csdg_adjudicate_claim()`. `csdg_export_assessment()` writes them to a directory, without models,
predictions, or row-level data.

```{r records}
as.data.table(chk1)[, .(gate, check, role, status)]
list.files(csdg_export_assessment(res1, tempdir()))
```

The numerical example of the article is worked through in
`vignette("claim_scoped_diagnostic_gates", package = "mlr3autoiml")`, and the
[function reference](https://stefancoors.de/mlr3autoiml/reference/index.html) documents all arguments and criteria.
