---
title: "The legacy AutoIML workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The legacy AutoIML workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

The `AutoIML` interface is the earlier workflow of `mlr3autoiml` and remains available for compatibility. Its gate
identifiers (G0A to G7B, with combined G3 and G6) and its pass/warn/fail/skip labels are heuristic workflow
diagnostics; they are separate from the CSDG gate registry (`csdg_gate_registry()`), the property statuses, and the
claim assessments. Printing a legacy `GateResult` emits a targeted deprecation warning. New analyses start from the
CSDG interface described in
[Checking what an explanation shows](https://stefancoors.de/mlr3autoiml/articles/exploratory_workflow.html)
(`vignette("exploratory_workflow", package = "mlr3autoiml")`).

The code in this article is shown but not run when the article is built, because each example fits several learners
with repeated resampling. To run it, install the suggested packages `mlr3learners`, `mlr3pipelines`, `ranger`,
`glmnet`, `xgboost`, and `rpart`.

## Example 1: classification for decision support

**Task:** `german_credit` (1,000 rows, 20 mixed-type features, binary `credit_risk`). The example runs the full
legacy gate path, including the combined G3 calibration and decision-utility gate and the G7A subgroup audit.

```{r classif-setup}
library(mlr3)
library(mlr3autoiml)
library(mlr3learners)
library(mlr3pipelines)

task = tsk("german_credit")
task$set_col_roles("personal_status_sex", add_to = "group")

learner = lrn("classif.rpart", predict_type = "prob", maxdepth = 6L)
```

The `ctx$claim` block is a pre-analysis declaration, not a model output.
The operative part is the claim scope, semantics, stakes, and any decision thresholds.
AutoIML can derive conservative defaults for non-use, prohibited interpretations,
and decision-policy wording from those declarations,
so you only need to spell them out when the analysis warrants narrower boundaries.

```{r classif-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "decision_support",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "decision_support",
  semantics = "within_support",
  stakes = "medium",
  claims = list(global = TRUE, local = FALSE, decision = TRUE),
  decision_spec = list(
    thresholds = seq(0.20, 0.60, by = 0.05),
    utility = list(tp = 1, fp = -0.5, fn = -1, tn = 0)
  )
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "personal_status_sex"
auto$ctx$alt_learners = list(
  featureless = lrn("classif.featureless", predict_type = "prob")
)
auto$ctx$multiplicity$transport_mode = "group_performance"
```

The measurement card needs only the part that the analysis cannot infer reliably, here the measurement level.
Gate 0B derives task- and pipeline-level notes for missingness and scoring automatically,
and asks for comparability or reliability evidence only when the measurement type makes those claims material.

```{r classif-run, results = "hide"}
auto$run()
```

```{r classif-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

With the operative claim card,
analysis-derived measurement notes, and subgroup audit in place,
the remaining warnings are substantive rather than setup-related:
calibration, net benefit in the declared threshold range,
and the interaction pattern that makes a simple one-feature story too thin.

```{r classif-g3, fig.width = 5.5, fig.height = 4.2, out.width = "49%", fig.show = "hold"}
auto$plot("g3_calibration")
auto$plot("g3_dca")
```

Gate 2 then shows why the package warns instead of presenting a single global
effect curve as the main story:
`amount` and `duration` dominate,
and they interact strongly.

```{r classif-g2, fig.width = 5.5, fig.height = 4.4, out.width = "49%", fig.show = "hold"}
auto$plot("g2_hstats", top_n = 3L)
auto$plot("g2_ale_2d", feature1 = "amount", feature2 = "duration", class_label = task$positive)
```

When Gate 2 regionalization is enabled
(`auto$ctx$structure$regionalize = TRUE`) and the interaction screen finds
materially heterogeneous effects, the GADGET-style views decompose the global
effect into interpretable regions and show the split rules that produced those
regions.
If no such regions are found, the Gate 2 evidence remains the
global effect plot, the interaction screen, and the top two-dimensional accumulated local effects (ALE) surface.
Optional `pint_enabled = TRUE` adds a permutation-based interaction screen
before regionalization.

```{r classif-g2-gadget}
if (!is.null(auto$result$gate_results$G2$artifacts$gadget_regions) &&
    nrow(auto$result$gate_results$G2$artifacts$gadget_regions) > 0L) {
  auto$plot("g2_gadget")
  auto$plot("g2_gadget_tree")
}
auto$plot("g2_pint")
```

```{r classif-export}
export_analysis_bundle(auto, dir = "bundle_german_credit", prefix = "german_credit")
```

## Example 2: regression for global insight

**Task:** `california_housing` (20,433 complete cases from the 20,640-row benchmark,
8 numeric features and 1 factor, continuous `median_house_value`).
We use a complete-case subset so `ranger`, lasso, and `xgboost` benchmark the same
design matrix. Correlated spatial predictors still trigger ALE selection in G2,
and the mixed learner family makes Gate 6 Rashomon diagnostics realistic.

```{r regr-setup}
task = tsk("california_housing")
task$filter(task$row_ids[stats::complete.cases(task$data())])

learner = lrn(
  "regr.ranger",
  num.trees = 200L,
  mtry.ratio = 0.5,
  min.node.size = 5L
)
```

The claim card fixes the semantics and scope before Gate 6 compares models.
Analysis-specific transport and preprocessing notes can remain implicit unless you need
to declare a narrower boundary than the task- and pipeline-level defaults.

```{r regr-autoiml}
auto = AutoIML$new(
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  purpose = "global_insight",
  seed = 42
)

auto$ctx$claim = make_claim_card(
  purpose = "global_insight",
  semantics = "within_support",
  stakes = "low",
  claims = list(global = TRUE, local = FALSE, decision = FALSE)
)

auto$ctx$measurement = make_measurement_card(
  level = "item"
)
auto$ctx$sensitive_features = "ocean_proximity"
auto$ctx$alt_learners = list(
  lasso = as_learner(po("encode") %>>% lrn("regr.glmnet", alpha = 1)),
  xgboost = as_learner(
    po("encode") %>>%
      lrn(
        "regr.xgboost",
        nrounds = 150L,
        eta = 0.05,
        max_depth = 6L,
        subsample = 0.8,
        colsample_bytree = 0.8,
        verbose = 0
      )
  ),
  tree = lrn("regr.rpart", maxdepth = 12L, cp = 1e-03),
  featureless = lrn("regr.featureless")
)
auto$ctx$structure$max_features = 8L
```

```{r regr-run, results = "hide"}
auto$run()
```

```{r regr-report}
knitr::kable(auto$report_card()[, .(gate_id, gate_name, status, summary)])
```

The effect plots focus on the housing variables that dominate the model story:
income and coastal location.
`ranger` keeps the global effects readable,
while Gate 6 benchmarks it against lasso, `xgboost`, and a single tree baseline.

```{r regr-g2, fig.width = 10, fig.height = 4.4, fig.alt = "ALE effects for median income and spatial coordinates"}
auto$plot("g2_effect", feature = c("median_income", "latitude", "longitude"))
```

The supporting views below add two complementary angles:
conditional SHAP (Shapley additive explanations) importance for the fitted forest,
and a Rashomon rank heatmap for the near-tie models in Gate 6.

```{r regr-shap-g6, fig.width = 5.8, fig.height = 4.6, out.width = "49%", fig.show = "hold"}
auto$plot("shap_importance", n_rows = 40L, sample_size = 15L, background_n = 40L, top_n = 8L)
auto$plot("g6_rank_heatmap", top_n = 8L)
```

```{r regr-export}
export_analysis_bundle(auto, dir = "bundle_california", prefix = "california")
```

## Gate overview of the legacy workflow

This table documents the legacy interface. The CSDG gate plan is returned by `csdg_gate_plan()`, and the CSDG gate
registry by `csdg_gate_registry()`.

| Gate | Triggered when | Evidence checks |
|------|---------------|-----------------|
| G0A  | always | Claim and semantics declaration; hard-stop on `causal_recourse` without identification |
| G0B  | always | Measurement readiness; reliability / invariance evidence for high-stakes |
| G1   | always | Cross-validated performance + calibration snapshot |
| G2   | always | Feature dependence · ALE vs partial dependence selection · pairwise interaction screening |
| G3   | `claims$decision = TRUE` | Calibration curve + decision curve analysis |
| G4   | `claims$local = TRUE` | SHAP local accuracy + perturbation sensitivity checks |
| G5   | always | Stability of narrated patterns under bootstrap resampling |
| G6   | high-resolution profile | Rashomon multiplicity + transport probes |
| G7A  | `sensitive_features` set | Subgroup performance + calibration audit |
| G7B  | user-facing + high-stakes | Human-factors (task-based evaluation) evidence |

## Compatibility status

The compatibility workflow uses a claim-first, gate-based architecture with
explicit semantics, direct gate evidence, and claim-dependent gate planning.
Default compatibility runs use the `high_resolution` profile;
`quick_start = TRUE` is available for rapid prototyping. Report cards and
guides retain the individual gate results and their claim-specific restrictions
instead of reducing them to a summary evidence score.
